---
title: "00_ncanda_demographics_cleaning2"
author: "Daniel Petrie"
date: "2025-04-01"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
#Packages
library(tidyverse) #Wrangling
library(dplyr) #Wrangling
library(stringr) #For str_replace
library(ggplot2)
library(psych)
library(BlandAltmanLeh)
#WD. Change whether through H: or on different sever
setwd("H:/Dan/ncanda/R")

#0-7
dem0_7 <- read.csv("H:/Raw/Datasets/NCANDA/followup_7y/NCANDA_RELEASE_7Y_REDCAP_MEASUREMENTS_V01/summaries/redcap/demographics.csv", header = TRUE)

#8
dem8 <- read.csv("H:/Raw/Datasets/NCANDA/followup_8y/NCANDA_RELEASE_8Y_REDCAP_MEASUREMENTS_V01/summaries/redcap/demographics.csv", header = TRUE)

#income file
income <- read.csv("H:/Raw/Datasets/NCANDA/followup_7y/NCANDA_RELEASE_7Y_REDCAP_MEASUREMENTS_V01/summaries/redcap/parentreport.csv")
```

Extracting basaeline demographic information. In particular, I want:

1. Income
2. Race/ethnicity

Note. I controlled for sex in all my analyses, so goal here is to just extract income/race and see if there are any differences between groups. I also 


```{r}
dem0_7_sub <- dem0_7 %>%
  dplyr::select(subject, visit, visit_age, hispanic, race, race_label) %>%
  filter(visit == "baseline")

income_sub <- income %>%
  dplyr::select(subject, visit, parentreport_age, 
                parentreport_sesp14, parentreport_sesp14_label,
                parentreport_sesp4, parentreport_sesp5) %>%
  filter(visit == "baseline")

eth_inc <- merge(income_sub, dem0_7_sub, by = c("subject", "visit"))
```


Getting a sense of distributions

```{r}
table(eth_inc$hispanic)
table(eth_inc$race)
table(eth_inc$race_label)

table(eth_inc$parentreport_sesp14_label)
table(eth_inc$parentreport_sesp14)

table(eth_inc$parentreport_sesp4)
table(eth_inc$parentreport_sesp5)
```

Dealing with ethnicity first. Here are race labels from data dictionary.

1, Native American/American Indian | 2, Asian | 3, Pacific Islander | 4, African-American/Black | 5, Caucasian/White | 6, Other (Specify) | 8, Invalid Entry | 9, No Response

This is ignoring hispanics.There are 97 self-dentified hispanics to incldue.

67 people put 6 down (some of which are hispanic). So need to decide what to do there. Probably just put into "other" for now.

I think it is reasonable to create 5 categories (asian/black/hispanic/white/other)

```{r}
#Create new column for race
eth_inc <- eth_inc %>%
  mutate(race_recode = case_when(
    hispanic == "Y" ~ 3,
    hispanic == "N" & race == 2 ~ 1,
    hispanic == "N" & race == 4 ~ 2,
    hispanic == "N" & race == 5 ~ 4,
    hispanic == "N" & race %in% c(1, 3, 6) ~ 5,
    TRUE ~ NA_real_  
  ),
  race_recode_labels = case_when(
    race_recode == 1 ~ "asian",
    race_recode == 2 ~ "black",
    race_recode == 3 ~ "hispanic",
    race_recode == 4 ~ "white",
    race_recode == 5 ~ "other",
    TRUE ~ NA_character_  
  ))


#eth_inc_6 <- eth_inc %>% filter(race == 6)
```

I also want to do the same with income. Below are the values

1, Less than $5,000 | 2, $5,000 through $11,999 | 3, $12,000 through $15,999 | 4, $16,000 through $24,999 | 5, $25,000 through $34,999 | 6, $35,000 through $49,999 | 7, $50,000 through $74,999 | 8, $75,000 through $99,999 | 9, $100,000 through $199,999 | 10, $200,000 and greater | 11, Don't know

Seems reasonable to chunk into 3/4 groups...
(1-6) < 50 k ~ 1
(7-8) >= 50 k and =< 100k ~ 2
(9-10) > 100k ~ 3
(11) Don't Know ~ 4

```{r}
eth_inc <- eth_inc %>%
  mutate(income_recode = case_when(
    parentreport_sesp14 %in% c(1:6) ~ 1,
    parentreport_sesp14 %in% c(7:8) ~ 2,
    parentreport_sesp14 %in% c(9:10) ~ 3,
    parentreport_sesp14 == 11 ~ 4,
    TRUE ~ NA_real_
  ),
  income_recode_labels = case_when(
    income_recode == 1  ~ "<50",
    income_recode == 2  ~ ">50<100",
    income_recode == 3  ~ ">100",
    income_recode == 4  ~ "dk", 
    TRUE ~ NA_character_
  ))
```


Household highest education I'll also look at.

1, High school diploma or equivalency (GED) | 2, Associate degree (junior college) | 3, Bachelor's degree (B.A., B.S.) | 4, Master's degree (M.A., M.S.) | 5, Doctorate (Ph.D., Ed.D.) | 6, Professional (MD, JD, DDS, etc.) | 7, Other | 8, None of the above (less than high school)

Grouping here feels odd to me. Let's try...
(1-2) High school/GED/AA ~ 1
(3) BA/BS ~ 2
(4-6) Post-graduate degree ~ 3
(7-8) Other/None ~ 4

```{r}
eth_inc <- eth_inc %>%
  mutate(high_edu = case_when(
    is.na(parentreport_sesp4) ~ parentreport_sesp5,  # If sesp4 is NA, take sesp5
    is.na(parentreport_sesp5) ~ parentreport_sesp4,  # If sesp5 is NA, take sesp4
    parentreport_sesp4 %in% c(7, 8) & parentreport_sesp5 %in% c(7, 8) ~ 8,  # If both are 7/8, take 8
    parentreport_sesp4 %in% c(7, 8) ~ parentreport_sesp5,  # If one is 7/8, take the other value
    parentreport_sesp5 %in% c(7, 8) ~ parentreport_sesp4,  
    TRUE ~ pmax(parentreport_sesp4, parentreport_sesp5, na.rm = TRUE)  # Otherwise, take the highest value
  ))

eth_inc <- eth_inc %>%
  mutate(
    edu = case_when(
      high_edu %in% c(1, 2) ~ 1,
      high_edu == 3 ~ 2,
      high_edu %in% c(4, 5, 6) ~ 3,
      high_edu %in% c(7, 8) ~ 4,
      TRUE ~ NA_real_
    ),
    edu_labels = case_when(
      edu == 1 ~ "High School/GED/AA",
      edu == 2 ~ "BA/BS",
      edu == 3 ~ "Post-Graduate Degree",
      edu == 4 ~ "Other/None",
      TRUE ~ NA_character_
    )
  )
```


Saving

```{r}
write.csv(eth_inc, file = "./clean_data/eth_inc.csv", row.names = FALSE)
```

