---
title: "07_fc_gam_clustering_cleaning"
output: html_document
date: "2025-10-10"
editor_options: 
  chunk_output_type: console
---

This script takes that cluster results and goes back to the individual/ROI level data and calculates unique cluster values for each subject. The script I am pulling from is "gam_cluster_cleaning".

End dataset should contain alcohol variables, FC variables that reflect cluster assignments, tissue iron just in case, and any other things I think might be relevant.

From here, I can run the analyses on the alcohol measures in a different script.

```{r}
#Packages
library(tidyverse) #Wrangling
library(dplyr) #Wrangling
library(stringr) #For str_replace
library(ggplot2)
library(psych)
library(lme4)
library(lmerTest)
library(GGally)
#library(pscl)
#library(glmmTMB)
library(ggeffects)
library(see)
#library(ggseg)
#library(ggsegGordon)
library(kableExtra)
library(clipr)
library(ggpubr)
#library(LNCDR)
library(gratia)
library(mgcv)
library(purrr)
library(readr)
library(emmeans)
#WD. Change whether through H: or on different sever
setwd("H:/Dan/ncanda/R/scripts")

#Gam model results
results_putamen <- read.csv("H:/Dan/ncanda/R/results/gam_models_putamen/gam_model_results_putamen.csv", 
                            header = TRUE)
results_caudate <- read.csv("H:/Dan/ncanda/R/results/gam_models_caudate/gam_model_results_caudate.csv", 
                            header = TRUE)
results_nacc <- read.csv("H:/Dan/ncanda/R/results/gam_models_nacc/gam_model_results_nacc.csv", 
                         header = TRUE)

#Clean seed data
dat_putamen <- read.csv("H:/Dan/ncanda/R/clean_data/full_putamen_seed_020425.csv", 
                        header = TRUE)
dat_caudate <- read.csv("H:/Dan/ncanda/R/clean_data/full_caudate_seed_020425.csv", 
                        header = TRUE)
dat_nacc <- read.csv("H:/Dan/ncanda/R/clean_data/full_nacc_seed_020425.csv", 
                     header = TRUE)

#Clusters for alcohol and FC groups
clusters <- read.csv("H:/Dan/ncanda/R/clean_data/clusters_alcohol.csv", 
                     header = TRUE)
gam_clusters <- read.csv("H:/Dan/ncanda/R/clean_data/gam_clusters.csv", 
                         header = TRUE)
```



This chunk does a lot

```{r}
#Creating sig column, filter by sig only, create -log p-value, change response to region, and remove "_harm".
results_putamen_sub <- results_putamen %>% 
  mutate(
    p_bonferroni = p.adjust(p_value_age, method = "bonferroni"),
    significant_fdr = p_fdr < 0.05, 
    significant_bonferroni = p_bonferroni < 0.05, 
    log_p_fdr = pmin(-log10(pmax(p_fdr, 1e-300)),6), #Check on this later, but no inf
    log_p_bonferroni = pmin(-log10(pmax(p_bonferroni, 1e-300)),6),
    region = str_remove(response, "_harm")) %>% 
  dplyr::select(region, everything()) %>% 
  filter(significant_fdr == TRUE | significant_bonferroni == TRUE)

#Creating sig column, filter by sig only, create -log p-value, change response to region, and remove "_harm".
results_caudate_sub <- results_caudate %>% 
  mutate(
    p_bonferroni = p.adjust(p_value_age, method = "bonferroni"),
    significant_fdr = p_fdr < 0.05, 
    significant_bonferroni = p_bonferroni < 0.05, 
    log_p_fdr = pmin(-log10(pmax(p_fdr, 1e-300)),6), #Check on this later, but no inf
    log_p_bonferroni = pmin(-log10(pmax(p_bonferroni, 1e-300)),6),
    region = str_remove(response, "_harm")) %>% 
  dplyr::select(region, everything()) %>% 
  filter(significant_fdr == TRUE | significant_bonferroni == TRUE)

#Creating sig column, filter by sig only, create -log p-value, change response to region, and remove "_harm".
results_nacc_sub <- results_nacc %>% 
  mutate(
    p_bonferroni = p.adjust(p_value_age, method = "bonferroni"),
    significant_fdr = p_fdr < 0.05, 
    significant_bonferroni = p_bonferroni < 0.05, 
    log_p_fdr = pmin(-log10(pmax(p_fdr, 1e-300)),6), #Check on this later, but no inf
    log_p_bonferroni = pmin(-log10(pmax(p_bonferroni, 1e-300)),6),
    region = str_remove(response, "_harm")) %>% 
  dplyr::select(region, everything()) %>% 
  filter(significant_fdr == TRUE | significant_bonferroni == TRUE)

#Combining into two main data frames. 
#One for significance, and other for growth models.

#Significance data
#Adding index column
results_putamen_sub <- results_putamen_sub %>% mutate(seed = "putamen")
results_caudate_sub <- results_caudate_sub %>% mutate(seed = "caudate")
results_nacc_sub <- results_nacc_sub %>% mutate(seed = "nacc")

#bind rows
results_combined <- bind_rows(results_putamen_sub, results_caudate_sub, results_nacc_sub)

#Filter by significant
#Note that old script also then filter by whether all three seeds are present.
results_significant <- results_combined %>%
  filter(significant_bonferroni) %>%
  #group_by(region) %>%
  #filter(all(c("putamen", "caudate", "nacc") %in% seed)) %>%
  ungroup()


#Growth model age plot data set. Note that I'll have to remove "_harm" if I want to plot on surface of brain.
#Adding k-mean cluster group
clusters_sub <- clusters %>%
  dplyr::select(subject, cluster)

dat_putamen <- merge(dat_putamen, clusters_sub, by = "subject")
dat_caudate <- merge(dat_caudate, clusters_sub, by = "subject")
dat_nacc <- merge(dat_nacc, clusters_sub, by = "subject")


#Define columns to retain that are unique for each seeded network.
putamen_responses <- results_significant %>%
  filter(seed == "putamen") %>%
  pull(response)
cols_to_keep_putamen <- c("subject", "age", "sex", "visit", "visit_numeric",
                  "cddr21", "cddr30", "cluster", "drinking_category_month",
                  #results_significant$response,
                  paste0(putamen_responses, "_z"))

caudate_responses <- results_significant %>%
  filter(seed == "caudate") %>%
  pull(response)
cols_to_keep_caudate <- c("subject", "age", "sex", "visit", "visit_numeric",
                  "cddr21", "cddr30", "cluster", "drinking_category_month",
                  #results_significant$response,
                  paste0(caudate_responses, "_z"))

nacc_responses <- results_significant %>%
  filter(seed == "nacc") %>%
  pull(response)
cols_to_keep_nacc <- c("subject", "age", "sex", "visit", "visit_numeric",
                  "cddr21", "cddr30", "cluster", "drinking_category_month",
                  #results_significant$response,
                  paste0(nacc_responses, "_z"))

# Step 2: Subset dat_putamen to keep only columns that match results_significant$region
dat_putamen_filtered <- dat_putamen %>%
  dplyr::select(any_of(cols_to_keep_putamen)) %>%
  mutate(seed = "putamen")

dat_caudate_filtered <- dat_caudate %>%
  dplyr::select(any_of(cols_to_keep_caudate)) %>%
  mutate(seed = "caudate")

dat_nacc_filtered <- dat_nacc %>%
  dplyr::select(any_of(cols_to_keep_nacc)) %>%
  mutate(seed = "nacc")

# Step 3: A combination 
#combined_long <- bind_rows(dat_putamen_filtered, dat_caudate_filtered, dat_nacc_filtered)

#I think keeping sep word be easier for now? Then binding right at the end.

# Reshape to long format
dat_putamen_filtered_long <- dat_putamen_filtered %>%
  pivot_longer(cols = ends_with("_z"), 
               names_to = "parcel", 
               values_to = "value") %>%
  mutate(
    drinking_category_month = factor(drinking_category_month,
                                     levels = c("no_low", "moderate", "heavy"))) %>%
  mutate(cluster_fac = factor(cluster, 
                              levels = c(1, 3, 2),  # New order (example)
                              labels = c("low_risk", "moderate_risk", "high_risk"))) %>%
  mutate(subject_fac = as.factor(subject),
         seed_fac = as.factor(seed)) %>%
  mutate(sex_fac = as.factor(sex))

dat_caudate_filtered_long <- dat_caudate_filtered %>%
  pivot_longer(cols = ends_with("_z"), 
               names_to = "parcel", 
               values_to = "value") %>%
  mutate(
    drinking_category_month = factor(drinking_category_month,
                                     levels = c("no_low", "moderate", "heavy"))) %>%
  mutate(cluster_fac = factor(cluster, 
                              levels = c(1, 3, 2),  # New order (example)
                              labels = c("low_risk", "moderate_risk", "high_risk"))) %>%
  mutate(subject_fac = as.factor(subject),
         seed_fac = as.factor(seed)) %>%
  mutate(sex_fac = as.factor(sex))

dat_nacc_filtered_long <- dat_nacc_filtered %>%
  pivot_longer(cols = ends_with("_z"), 
               names_to = "parcel", 
               values_to = "value") %>%
  mutate(
    drinking_category_month = factor(drinking_category_month,
                                     levels = c("no_low", "moderate", "heavy"))) %>%
  mutate(cluster_fac = factor(cluster, 
                              levels = c(1, 3, 2),  # New order (example)
                              labels = c("low_risk", "moderate_risk", "high_risk"))) %>%
  mutate(subject_fac = as.factor(subject),
         seed_fac = as.factor(seed)) %>%
  mutate(sex_fac = as.factor(sex))
```

OK. I need to add clusters for the gam fits.

```{r}
dat_putamen_filtered_long_t <- dat_putamen_filtered_long %>%
  mutate(parcel_nosuffix = str_remove(parcel, "_z")) %>%
  left_join(gam_clusters, 
            by = c("seed" = "Seed", "parcel_nosuffix" = "Response")) %>%
  rename(gam_clusters = Cluster) %>%
  select(-parcel_nosuffix, -Count)  # optional: drop the helper column

dat_caudate_filtered_long_t <- dat_caudate_filtered_long %>%
  mutate(parcel_nosuffix = str_remove(parcel, "_z")) %>%
  left_join(gam_clusters, 
            by = c("seed" = "Seed", "parcel_nosuffix" = "Response")) %>%
  rename(gam_clusters = Cluster) %>%
  select(-parcel_nosuffix, -Count)  # optional: drop the helper column

dat_nacc_filtered_long_t <- dat_nacc_filtered_long %>%
  mutate(parcel_nosuffix = str_remove(parcel, "_z")) %>%
  left_join(gam_clusters, 
            by = c("seed" = "Seed", "parcel_nosuffix" = "Response")) %>%
  rename(gam_clusters = Cluster) %>%
  select(-parcel_nosuffix, -Count)  # optional: drop the helper column
```

Ok. Now I have clusters assigned to the data...

Next step is to try and get trajectories averages by cluster now. Should create 6 new columns per subject per visit essentially...

```{r}
dat_putamen_filtered_long_t_avg <- dat_putamen_filtered_long_t %>%
  group_by(subject, visit, seed, gam_clusters) %>%
  summarise(mean_value = mean(value, na.rm = TRUE), .groups = "drop") %>%
  unite("seed_cluster", seed, gam_clusters, sep = "_cluster") %>%
  pivot_wider(names_from = seed_cluster, values_from = mean_value)

dat_caudate_filtered_long_t_avg <- dat_caudate_filtered_long_t %>%
  group_by(subject, visit, seed, gam_clusters) %>%
  summarise(mean_value = mean(value, na.rm = TRUE), .groups = "drop") %>%
  unite("seed_cluster", seed, gam_clusters, sep = "_cluster") %>%
  pivot_wider(names_from = seed_cluster, values_from = mean_value)

dat_nacc_filtered_long_t_avg <- dat_nacc_filtered_long_t %>%
  group_by(subject, visit, seed, gam_clusters) %>%
  summarise(mean_value = mean(value, na.rm = TRUE), .groups = "drop") %>%
  unite("seed_cluster", seed, gam_clusters, sep = "_cluster") %>%
  pivot_wider(names_from = seed_cluster, values_from = mean_value)
```

Adding back to subject-level data.

```{r}
dat_putamen_final <- dat_putamen_filtered_long_t %>%
  select(subject, age, sex, visit, visit_numeric, cddr21, cddr30, drinking_category_month, cluster_fac, subject_fac, seed_fac, sex_fac) %>%
  distinct() %>%
  left_join(dat_putamen_filtered_long_t_avg, by = c("subject", "visit"))

dat_caudate_final <- dat_caudate_filtered_long_t %>%
  select(subject, age, sex, visit, visit_numeric, cddr21, cddr30, drinking_category_month, cluster_fac, subject_fac, seed_fac, sex_fac) %>%
  distinct() %>%
  left_join(dat_caudate_filtered_long_t_avg, by = c("subject", "visit"))

dat_nacc_final <- dat_nacc_filtered_long_t %>%
  select(subject, age, sex, visit, visit_numeric, cddr21, cddr30, drinking_category_month, cluster_fac, subject_fac, seed_fac, sex_fac) %>%
  distinct() %>%
  left_join(dat_nacc_filtered_long_t_avg, by = c("subject", "visit"))
```

Using full join

```{r}
dat_combined_final <- dat_putamen_final %>%
  left_join(dat_caudate_final %>% select(-age, -sex, -visit_numeric, -cddr21, -cddr30, -drinking_category_month,
                                         -cluster_fac, -subject_fac, -seed_fac, -sex_fac),
            by = c("subject", "visit")) %>%
  left_join(dat_nacc_final %>% select(-age, -sex, -visit_numeric, -cddr21, -cddr30, -drinking_category_month,
                                      -cluster_fac, -subject_fac, -seed_fac, -sex_fac),
            by = c("subject", "visit"))
```

Longer

```{r}
dat_combined_final_long <- dat_combined_final %>%
  pivot_longer(
    cols = c(putamen_cluster1, putamen_cluster2,
             caudate_cluster1, caudate_cluster2,
             nacc_cluster1, nacc_cluster2),
    names_to = "seed_cluster",
    values_to = "mean_value"
  )
```


Saving these files.

```{r Saving}
write.csv(dat_combined_final, file = "H:/Dan/ncanda/R/clean_data/fc_clusters_alcohol_101025.csv", 
          row.names = FALSE)
write.csv(dat_combined_final_long, file = "H:/Dan/ncanda/R/clean_data/fc_clusters_alcohol_long_101025.csv", 
          row.names = FALSE)
```



Lets hold off on iron for now.

```{r}
#Adding iron
dat_combined_final_iron <- dat_combined_final %>% 
  left_join(
  dat_putamen %>% select(subject, visit, containssss("iron_harm_z")),
  by = c("subject", "visit")
)
```

