---
title: "09_binge_specificity"
output: html_document
date: "2026-03-16"
editor_options: 
  chunk_output_type: console
---

Libraries, wd(), and data

```{r Global}
#Packages
library(tidyverse) #Wrangling
library(dplyr) #Wrangling
library(stringr) #For str_replace
library(ggplot2)
library(psych)
library(lme4)
library(lmerTest)
library(GGally)
#library(pscl)
#library(glmmTMB)
library(ggeffects)
library(see)
#library(ggseg)
#library(ggsegGordon)
library(kableExtra)
library(clipr)
library(ggpubr)
#library(LNCDR)
library(gratia)
library(mgcv)
library(purrr)
library(readr)
library(emmeans)
library(data.table)
library(ggridges)
library(ggnewscale)
#WD. Change whether through H: or on different sever
setwd("H:/Dan/ncanda/R/")

#Clean data
dat_long <- read.csv("./clean_data/fc_clusters_alcohol_long_101025.csv", header = TRUE)
dat <- read.csv("./clean_data/fc_clusters_alcohol_101025.csv", header = TRUE)
```

# Colors I am using for now. Might have to switch to some colorblind friendly pallete.

```{r Color pallete}
  scale_color_manual(values = c("putamen_cluster1" = "#1b9e77",
                                "putamen_cluster2" = "#1b3877ff",
                                "caudate_cluster1" = "#7570b3",
                                "caudate_cluster2" = "#7500b3ff",
                                "nacc_cluster1" = "#d95f02",
                                "nacc_cluster2" = "#e6ab02")) +# putamen_cluster1
scale_color_gradient(low = "#cceae7", high = "#1b9e77")

# putamen_cluster2
scale_color_gradient(low = "#cdd6e8", high = "#1b3877ff")

# caudate_cluster1
scale_color_gradient(low = "#e3def2", high = "#7570b3")

# caudate_cluster2
scale_color_gradient(low = "#e4d6f2", high = "#7500b3ff")

# nacc_cluster1
scale_color_gradient(low = "#f6d8c6", high = "#d95f02")

# nacc_cluster2
scale_color_gradient(low = "#f8edcc", high = "#e6ab02")

```


# Primary analyses for alcohol use behavior

## TVEM analyses

TVEM to predict trajectories of past year days drinking based on corticostriatal FC. I will fit 6 models and use a Bonferonni correction. For any significant time varying effects, I will further examine using derivatives and plotting the associations.

```{r}
by_vars <- c(
  "putamen_cluster1",
  "putamen_cluster2",
  "caudate_cluster1",
  "caudate_cluster2",
  "nacc_cluster1",
  "nacc_cluster2"
)

# Storage lists
fits <- list()
model_warnings <- list()

for (v in by_vars) {

  # build formula
  formula <- as.formula(
    paste0(
      "cddr30 ~ 1 + sex_fac + 
       s(age, k = 5) + 
       s(age, by = ", v, ", k = 5)"
    )
  )

  # fit model
  fit <- gamm(
    formula,
    random = list(subject = ~1, subject = ~0 + age),
    data = dat,
    method = "REML",
    na.action = na.exclude
  )

  # store model
  fits[[v]] <- fit

  # capture warnings produced during summary()
  w <- NULL
  invisible(
    withCallingHandlers(
      summary(fit$gam),
      warning = function(wrn) {
        w <<- c(w, conditionMessage(wrn))
        invokeRestart("muffleWarning")
      }
    )
  )

  model_warnings[[v]] <- w
}

# Optional: print GAM summaries
lapply(fits, function(x) summary(x$gam))
lapply(fits, function(x) summary(x$lme))

lapply(fits, function(x) gam.vcomp(x$gam))

model_warnings
```

We can safely ignore these warnings. They are a function of the numeric-by smooths and are benign.

```{r}
# Extract only "by" smooth p-values across all models
by_smooth_pvals <- map_df(seq_along(fits), function(i) {
  s <- summary(fits[[i]]$gam)
  
  # Keep only smooths where the term name contains "by=" so : esssentially
  smooth_table <- as.data.frame(s$s.table) %>%
    mutate(term = rownames(s$s.table),
           model = i) %>%
    filter(grepl(":", term)) %>%
    select(model, term, edf = edf, F, p_value = `p-value`)
  
  smooth_table
})

# Apply FDR correction across these "by" smooths
by_smooth_pvals <- by_smooth_pvals %>%
  mutate(p_value_fdr = p.adjust(p_value, method = "bonferroni"))

by_smooth_pvals

#putamen <- fits[[2]]$gam$pterms

#pred_growth <- ggpredict(putamen, terms = c("age[17:30]", 
#                                           "putamen_cluster2[-2,-1,0,1,2]"))

#plot(pred_growth)
```

All not significant. No need to plot.

## TVEM figure(s)

```{r TVEM figures BINGE}
#Putamen cluster1
pred_growth <-  ggpredict(fits$putamen_cluster1, terms = c("age[12.04:29.3]", 
                                           "putamen_cluster1[-2,-1,0,1,2]"))

plot(pred_growth, show_data = F)
#Extract derivatives
d <- derivatives(fits$putamen_cluster1,
                 interval = "simultaneous",
                 n_sim = 10000,
                 n = 200)
#Plot derivatives
draw(d)
#21.14 years
#Calculate regions of significance, where CI != 0
d <- d %>% mutate(sig = !(0>.lower_ci & 0<.upper_ci))

#Nice figure
p3 <- ggplot(data = pred_growth, 
             aes(x = x, 
                 y = predicted,
                 group = group,
                 colour = as.numeric(as.character(group)))) +
  geom_line() +
  scale_color_gradient(low = "#cceae7", high = "#1b9e77") +
  theme_modern() +
  geom_vline(xintercept = 22.79, 
             linetype = "dashed", 
             color = "black", 
             linewidth = 0.5) +
  labs(x = "Age (years)",
       y = "Days of Alcohol Use Past Year",
       color = "Putamen Cluster 1\n(z-scored)") +
    theme(
    axis.title.x = element_text(size = 8, margin = margin(t=3)),
    axis.title.y = element_text(size = 8, margin = margin(r=3)),
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 7),
    legend.key.size = unit(0.3, "cm")
  ) +
  guides(colour = guide_legend(reverse = TRUE))
p3

ggsave(p3, 
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_binge/tvem_put1.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 90, 
       height = 60)
#
#
#Putamen cluster2
#
#
pred_growth <-  ggpredict(fits$putamen_cluster2, terms = c("age[12.04:29.3]", 
                                           "putamen_cluster2[-2,-1,0,1,2]"))

plot(pred_growth, show_data = TRUE)
#Extract derivatives
d <- derivatives(fits$putamen_cluster2,
                 interval = "simultaneous",
                 n_sim = 10000,
                 n = 200)
#Plot derivatives
draw(d)
#21.14 years
#Calculate regions of significance, where CI != 0
d <- d %>% mutate(sig = !(0>.lower_ci & 0<.upper_ci))

#Nice figure
p3 <- ggplot(data = pred_growth, 
             aes(x = x, 
                 y = predicted,
                 group = group,
                 colour = as.numeric(as.character(group)))) +
  geom_line() +
  scale_color_gradient(low = "#cdd6e8", high = "#1b3877ff") +
  theme_modern() +
  #geom_vline(xintercept = 22.79, 
  #           linetype = "dashed", 
  #           color = "black", 
  #           linewidth = 0.5) +
  labs(x = "Age (years)",
       y = "Days of Alcohol Use Past Year",
       color = "Putamen Cluster 2\n(z-scored)") +
    theme(
    axis.title.x = element_text(size = 8, margin = margin(t=3)),
    axis.title.y = element_text(size = 8, margin = margin(r=3)),
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 7),
    legend.key.size = unit(0.3, "cm")
  ) +
  guides(colour = guide_legend(reverse = TRUE))
p3

ggsave(p3, 
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_binge/tvem_put2.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 90, 
       height = 60)

#
#
#caud cluster1
#
#
pred_growth <-  ggpredict(fits$caudate_cluster1, terms = c("age[12.04:29.3]", 
                                           "caudate_cluster1[-2,-1,0,1,2]"))

plot(pred_growth, show_data = TRUE)
#Extract derivatives
d <- derivatives(fits$caudate_cluster1,
                 interval = "simultaneous",
                 n_sim = 10000,
                 n = 200)
#Plot derivatives
draw(d)
#21.14 years
#Calculate regions of significance, where CI != 0
d <- d %>% mutate(sig = !(0>.lower_ci & 0<.upper_ci))

#Nice figure
p3 <- ggplot(data = pred_growth, 
             aes(x = x, 
                 y = predicted,
                 group = group,
                 colour = as.numeric(as.character(group)))) +
  geom_line() +
  scale_color_gradient(low = "#e3def2", high = "#7570b3") +
  theme_modern() +
  #geom_vline(xintercept = 22.79, 
  #           linetype = "dashed", 
  #           color = "black", 
  #           linewidth = 0.5) +
  labs(x = "Age (years)",
       y = "Days of Alcohol Use Past Year",
       color = "Caudate Cluster 1\n(z-scored)") +
    theme(
    axis.title.x = element_text(size = 8, margin = margin(t=3)),
    axis.title.y = element_text(size = 8, margin = margin(r=3)),
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 7),
    legend.key.size = unit(0.3, "cm")
  ) +
  guides(colour = guide_legend(reverse = TRUE))
p3

ggsave(p3, 
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_binge/tvem_caud1.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 90, 
       height = 60)

#
#
#caud cluster2
#
#
pred_growth <-  ggpredict(fits$caudate_cluster2, terms = c("age[12.04:29.3]", 
                                           "caudate_cluster2[-2,-1,0,1,2]"))

plot(pred_growth, show_data = TRUE)
#Extract derivatives
d <- derivatives(fits$caudate_cluster2,
                 interval = "simultaneous",
                 n_sim = 10000,
                 n = 200)
#Plot derivatives
draw(d)
#21.14 years
#Calculate regions of significance, where CI != 0
d <- d %>% mutate(sig = !(0>.lower_ci & 0<.upper_ci))

#Nice figure
p3 <- ggplot(data = pred_growth, 
             aes(x = x, 
                 y = predicted,
                 group = group,
                 colour = as.numeric(as.character(group)))) +
  geom_line() +
  scale_color_gradient(low = "#e4d6f2", high = "#7500b3ff") +
  theme_modern() +
  #geom_vline(xintercept = 22.79, 
  #           linetype = "dashed", 
  #           color = "black", 
  #           linewidth = 0.5) +
  labs(x = "Age (years)",
       y = "Days of Alcohol Use Past Year",
       color = "Caudate Cluster 2\n(z-scored)") +
    theme(
    axis.title.x = element_text(size = 8, margin = margin(t=3)),
    axis.title.y = element_text(size = 8, margin = margin(r=3)),
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 7),
    legend.key.size = unit(0.3, "cm")
  ) +
  guides(colour = guide_legend(reverse = TRUE))
p3

ggsave(p3, 
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_binge/tvem_caud2.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 90, 
       height = 60)


#
#
#nacc cluster1
#
#
pred_growth <-  ggpredict(fits$nacc_cluster1, terms = c("age[12.04:29.3]", 
                                           "nacc_cluster1[-2,-1,0,1,2]"))

plot(pred_growth, show_data = TRUE)
#Extract derivatives
d <- derivatives(fits$nacc_cluster1,
                 interval = "simultaneous",
                 n_sim = 10000,
                 n = 200)
#Plot derivatives
draw(d)
#21.14 years
#Calculate regions of significance, where CI != 0
d <- d %>% mutate(sig = !(0>.lower_ci & 0<.upper_ci))

#Nice figure
p3 <- ggplot(data = pred_growth, 
             aes(x = x, 
                 y = predicted,
                 group = group,
                 colour = as.numeric(as.character(group)))) +
  geom_line() +
  scale_color_gradient(low = "#f6d8c6", high = "#d95f02") +
  theme_modern() +
  #geom_vline(xintercept = 22.79, 
  #           linetype = "dashed", 
  #           color = "black", 
  #           linewidth = 0.5) +
  labs(x = "Age (years)",
       y = "Days of Alcohol Use Past Year",
       color = "NAcc Cluster 1\n(z-scored)") +
    theme(
    axis.title.x = element_text(size = 8, margin = margin(t=3)),
    axis.title.y = element_text(size = 8, margin = margin(r=3)),
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 7),
    legend.key.size = unit(0.3, "cm")
  ) +
  guides(colour = guide_legend(reverse = TRUE))
p3

ggsave(p3, 
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_binge/tvem_nacc1.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 90, 
       height = 60)


#
#
#nacc cluster1
#
#
pred_growth <-  ggpredict(fits$nacc_cluster2, terms = c("age[12.04:29.3]", 
                                           "nacc_cluster2[-2,-1,0,1,2]"))

plot(pred_growth, show_data = TRUE)
#Extract derivatives
d <- derivatives(fits$nacc_cluster2,
                 interval = "simultaneous",
                 n_sim = 10000,
                 n = 200)
#Plot derivatives
draw(d)
#21.14 years
#Calculate regions of significance, where CI != 0
d <- d %>% mutate(sig = !(0>.lower_ci & 0<.upper_ci))

#Nice figure
p3 <- ggplot(data = pred_growth, 
             aes(x = x, 
                 y = predicted,
                 group = group,
                 colour = as.numeric(as.character(group)))) +
  geom_line() +
  scale_color_gradient(low = "#f8edcc", high = "#e6ab02") +
  theme_modern() +
  #geom_vline(xintercept = 22.79, 
  #           linetype = "dashed", 
  #           color = "black", 
  #           linewidth = 0.5) +
  labs(x = "Age (years)",
       y = "Days of Alcohol Use Past Year",
       color = "NAcc Cluster 2\n(z-scored)") +
    theme(
    axis.title.x = element_text(size = 8, margin = margin(t=3)),
    axis.title.y = element_text(size = 8, margin = margin(r=3)),
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 7),
    legend.key.size = unit(0.3, "cm")
  ) +
  guides(colour = guide_legend(reverse = TRUE))
p3

ggsave(p3, 
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_binge/tvem_nacc2.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 90, 
       height = 60)
```

### TVEM raw data plots. These are for supplementary fig if I end up adding.

```{r TVEM w/ Raw data plot}
pred_growth_p4 <-  ggpredict(fits$putamen_cluster2, terms = c("age[12.04:29.3]", 
                                           "putamen_cluster2[-2,-1,0,1,2]"))
p4 <- ggplot() +
  geom_line(data = pred_growth_p4, 
             aes(x = x, 
                 y = predicted,
                 group = group,
                 colour = as.numeric(as.character(group))),
            show.legend = FALSE) +
  scale_color_gradient(low = "#cceae7", high = "#1b9e77") +
  new_scale_color() +
  new_scale_fill() +
  geom_point(data = dat, 
             aes(x = age, 
                 y = cddr30),
             size = 0.05,
             alpha = 0.2,
             color = "grey60") + 
  #geom_line(data = full, 
  #          aes(x = age, 
  #              y = cic_pos_dorsal_put_harm, 
  #              color = mb_resid_z, 
  #              group = id),
  #          linewidth = 0.5) +
  theme_modern() +
  geom_vline(xintercept = 22.79, 
             linetype = "dashed", 
             color = "black", 
             linewidth = 0.5) +
  labs(x = "Age (years)",
       y = "Days of Alcohol Use Past Year",
       color = "Putamen FC\n(z-scored)") +
    theme(
    axis.title.x = element_text(size = 8, margin = margin(t=3)),
    axis.title.y = element_text(size = 8, margin = margin(r=3)),
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 7),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 7),
    legend.key.size = unit(0.3, "cm")
  ) +
  guides(colour = guide_legend(reverse = TRUE))
  
p4

ggsave(p4, 
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_binge/tvem_put1_raw_PRESENTATION_FIGURE.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 90, 
       height = 60)
```



# Drinking initiation analyses

Idea here is to control for age effects of FC and predict changes in connectivity based on functional connectivity cluster and drinking phase.

```{r Drinking initiation main analysis}
#Drinking onset? Before/after/never drank
dat_long_ba <- dat_long %>%
  group_by(subject) %>%
  mutate(
    drinking_onset = ifelse(any(cddr30 > 0), min(age[cddr30 > 0], na.rm = TRUE), NA),
    drinking_phase = case_when(
      is.na(drinking_onset) ~ "Never",   # Never drank
      age < drinking_onset  ~ "Before",  # Before first drinking report
      age >= drinking_onset ~ "After"    # After first drinking report
    )
  ) %>%
  ungroup() %>%
  mutate(drinking_phase = factor(drinking_phase, levels = c("Before", "After", "Never")),
         seed_fac = factor(seed_fac),
         subject_fac = factor(subject_fac))


#Excplicity testing via anova-like approach
dat_long_ba <- dat_long_ba %>%
  mutate(seed_cluster_fac = factor(seed_cluster,
                                   levels = c("nacc_cluster1", "nacc_cluster2", 
                                              "caudate_cluster1", "caudate_cluster2", 
                                              "putamen_cluster1", "putamen_cluster2")))


please_lord <- bam(mean_value ~ 1 + sex_fac  +  seed_cluster_fac*drinking_phase +
                                s(age, k = 5) + 
                                s(subject_fac, bs = 're') +
                                s(subject_fac, age, bs = 're'), 
                   data = dat_long_ba, 
                   na.action = na.exclude, 
                   method = "fREML")
summary(please_lord)

plot(ggpredict(please_lord, terms = c("drinking_phase", 
                                      "seed_cluster_fac")))


#Probing the intereactions in a few different ways
emm_results <- emmeans(please_lord, ~ drinking_phase | seed_cluster_fac)

pairs(emm_results)

contrasts <- contrast(emm_results, method = "pairwise", adjust = "none")
contrasts

#Just doing adjustment manually.
contrasts_dat <- as.data.frame(contrasts)
contrasts_dat$p_adj <- p.adjust(contrasts_dat$p.value, method = "bonferroni")

#making sig column for easy viewing.
contrasts_dat$sig <- contrasts_dat$p_adj < 0.05
contrasts_dat
#Plotting contrasts
plot(emm_results, horizontal = TRUE)
plot(contrasts)

q <- as.data.frame(emm_results)
```

Drinking initiation is associated with increased FC in NAcc cluster 1 and Caudate cluster 2. 



## Drinking initiation figures

### Main figure showing results from main initiation model

```{r Initiation Figure all ROIs}
seed_phase <- ggplot(q %>% 
                       filter(drinking_phase %in% c("Before", "After", "Never")), 
                     aes(x = seed_cluster_fac, 
                         y = emmean, 
                         group = drinking_phase,
                         color = seed_cluster_fac)) + 
  geom_point(position = position_dodge(width = 0.7), size = 1) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.2, 
                position = position_dodge(width = 0.7)) +
  scale_color_manual(values = c("putamen_cluster1" = "#1b9e77",
                                "putamen_cluster2" = "#1b3877ff",
                                "caudate_cluster1" = "#7570b3",
                                "caudate_cluster2" = "#7500b3ff",
                                "nacc_cluster1" = "#d95f02",
                                "nacc_cluster2" = "#e6ab02"
                                )) +
  scale_x_discrete(labels = rep(c("Before    After   Never"), 6)) +
  annotate("text", x = 1:6, y = .20, 
           label = c("NAcc Cluster 1", "NAcc Cluster 2", 
                     "Caudate Cluster 1", "Caudate Cluster 2",
                     "Putamen Cluster 1", "Putamen Cluster 2"),
           vjust = -0.5, hjust = 0.5, size = 2) +
  labs(x = "Drinking Phase", y = "Estimated Mean") +
  theme_modern() +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    legend.position = "none"
    )
seed_phase

ggsave(plot = seed_phase,
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_binge/seed_phase.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 160, 
       height = 60)


#I also want to isolate figures that I will carry over. Putamen 1
seed_phase <- ggplot(q %>% 
                       filter(drinking_phase %in% c("Before", "After", "Never")) %>%
                       filter(seed_cluster_fac == "putamen_cluster1"), 
                     aes(x = seed_cluster_fac, 
                         y = emmean, 
                         group = drinking_phase,
                         color = seed_cluster_fac)) + 
  geom_point(position = position_dodge(width = 0.7), size = 1) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.2, 
                position = position_dodge(width = 0.7)) +
  scale_color_manual(values = c("putamen_cluster1" = "#1b9e77")) +
  scale_y_continuous(limits = c(-0.2, 0.2),
                     breaks = seq(-0.2,0.2, by = 0.1)) +
  #scale_x_discrete(labels = rep(c("Before    After   Never"), 1)) +
  #annotate("text", x = 0, y = .20, 
  #         label = c("Putamen Cluster 1"),
  #         vjust = -0.5, hjust = 0.5, size = 2) +
  labs(x = "", y = "") +
  theme_modern() +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    legend.position = "none"
    )
seed_phase

ggsave(plot = seed_phase,
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_binge/putamen1_init.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 36, 
       height = 60)

#I also want to isolate figures that I will carry over. caudate 2
seed_phase <- ggplot(q %>% 
                       filter(drinking_phase %in% c("Before", "After", "Never")) %>%
                       filter(seed_cluster_fac == "caudate_cluster2"), 
                     aes(x = seed_cluster_fac, 
                         y = emmean, 
                         group = drinking_phase,
                         color = seed_cluster_fac)) + 
  geom_point(position = position_dodge(width = 0.7), size = 1) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.2, 
                position = position_dodge(width = 0.7)) +
  scale_color_manual(values = c("caudate_cluster2" = "#7500b3ff")) +
  scale_y_continuous(limits = c(-0.2, 0.2),
                     breaks = seq(-0.2,0.2, by = 0.1)) +
  #scale_x_discrete(labels = rep(c("Before    After   Never"), 1)) +
  #annotate("text", x = 0, y = .20, 
  #         label = c("Putamen Cluster 1"),
  #         vjust = -0.5, hjust = 0.5, size = 2) +
  labs(x = "", y = "") +
  theme_modern() +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    legend.position = "none"
    )
seed_phase

ggsave(plot = seed_phase,
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_binge/caudate2_init.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 36, 
       height = 60)

#I also want to isolate figures that I will carry over. nacc 1
seed_phase <- ggplot(q %>% 
                       filter(drinking_phase %in% c("Before", "After", "Never")) %>%
                       filter(seed_cluster_fac == "nacc_cluster1"), 
                     aes(x = seed_cluster_fac, 
                         y = emmean, 
                         group = drinking_phase,
                         color = seed_cluster_fac)) + 
  geom_point(position = position_dodge(width = 0.7), size = 1) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.2, 
                position = position_dodge(width = 0.7)) +
  scale_color_manual(values = c("nacc_cluster1" = "#d95f02")) +
  scale_y_continuous(limits = c(-0.2, 0.2),
                     breaks = seq(-0.2,0.2, by = 0.1)) +
  #scale_x_discrete(labels = rep(c("Before    After   Never"), 1)) +
  #annotate("text", x = 0, y = .20, 
  #         label = c("Putamen Cluster 1"),
  #         vjust = -0.5, hjust = 0.5, size = 2) +
  labs(x = "", y = "") +
  theme_modern() +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    legend.position = "none"
    )
seed_phase

ggsave(plot = seed_phase,
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_binge/nacc1_init.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 36, 
       height = 60)
```


### Analyses and Figure showing growth model colored by drinking onset (nacc)


```{r}
dat_long_ba$time_since_onset <- dat_long_ba$age - dat_long_ba$drinking_onset

#Creating new dataframe
dat_pred_nacc1 <- dat_long_ba %>%
  filter(seed_cluster == "nacc_cluster1") 

#Seeing if additional grouping variable works
dat_pred_nacc1 <- dat_pred_nacc1 %>%
  mutate(drinking_group = ifelse(drinking_phase == "Never", "Never", "Initiated"),
         drinking_group = factor(drinking_group, levels = c("Initiated", "Never")))

#Oredered?
dat_pred_nacc1$drinking_group_ord <- factor(dat_pred_nacc1$drinking_group, 
                             levels = c("Never", "Initiated"), 
                             ordered = TRUE)

test <- bam(mean_value ~ 1 + sex_fac + drinking_group_ord +
              s(age, k = 5) +
              s(age, by = drinking_group_ord, k = 5) +
              s(subject_fac, bs = 're') +
              s(age, subject_fac, bs = 're'), 
            data = dat_pred_nacc1, 
            na.action = na.exclude, 
            method = "fREML")

summary(test)
#Function to calculate difference smooths
#test_diff <- difference_smooths(test, select = c("s(age)", "s(age):drinking_group_ordInitiated"), partial_match = FALSE)

#draw(test_diff)

#Gratia plot
draw(test)
#mgcv plot
plot(test, pages = 1)

plot(ggpredict(test, terms = c("age")))
plot(ggpredict(test, terms = c("drinking_group_ord")))
plot(ggpredict(test, terms = c("age", "drinking_group_ord")), show_ci = FALSE)


# Add predicted values to your data

dat_pred_nacc1$predicted <-  predict(test, newdata = dat_pred_nacc1, type = "response")




#If I want to color lines accordingly, I need to create average age by drinking-phase.
cutoff_age <- dat_pred_nacc1 %>%
  filter(drinking_phase == "After") %>%
  group_by(subject) %>%
  summarize(min_after_age = min(age), .groups = "drop") %>%
  summarize(avg_cutoff = mean(min_after_age, na.rm = TRUE)) %>%
  pull(avg_cutoff)

dat_pred_nacc1 <- dat_pred_nacc1 %>%
  mutate(phase_age = case_when(
    drinking_phase == "Never" ~ "Never",
    age < cutoff_age ~ "Before",
    age >= cutoff_age ~ "After"
  ))

nacc1 <- ggplot(dat_pred_nacc1, aes(x = age, y = predicted, group = subject, color = drinking_phase)) +
  #geom_point(alpha = 0.2, size = 0.05) +
  geom_line(alpha = 0.2, size = 0.1) +
      scale_color_manual(values = c(
    "After" = "#d95f02ff",
    "Before" = "black", 
    "Never" = "black"))   +   
    new_scale_color() +
  new_scale_fill() +
  geom_smooth(aes(group = phase_age, color = phase_age), size = 0.3, show.legend = FALSE) +
  labs(y = "", x = "") +
    scale_color_manual(values = c(
    "After" = "#d95f02ff",  
    "Before" = "black", 
    "Never" = "black"
  )) +
    facet_wrap(~drinking_group, nrow = 1) + 
 # geom_vline(xintercept = cutoff_age, linetype = "dashed", size = 0.2) +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    legend.position = "none",
    strip.text = element_blank(),
    panel.background = element_blank(),   # remove gray panel
    panel.grid = element_blank(),         # remove grid lines
    plot.background = element_blank()     # remove outer gray
  )

nacc1

ggsave(filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_binge/NAcc_pred_ban_plot.pdf", 
      plot = nacc1,
       device = "pdf", 
       dpi = 300, 
      units = "mm",
       width = 90, 
       height = 60)
```

### Figure showing growth model colored by drinking onset (caudate)

```{r}
# Add predicted values to your data
dat_pred_caud2 <- dat_long_ba %>%
  filter(seed_cluster == "caudate_cluster2") 

#Seeing if additional grouping variable works
dat_pred_caud2 <- dat_pred_caud2 %>%
  mutate(drinking_group = ifelse(drinking_phase == "Never", "Never", "Initiated"),
         drinking_group = factor(drinking_group, levels = c("Initiated", "Never")))

#Oredered?
dat_pred_caud2$drinking_group_ord <- factor(dat_pred_caud2$drinking_group, 
                             levels = c("Never", "Initiated"), 
                             ordered = TRUE)

#Seeing if there are actual trajectory differences
test2 <- bam(mean_value ~ 1 + sex_fac  +  drinking_group_ord +
              s(age, k = 5) +
              s(age, by = drinking_group_ord, k = 5) +
              s(subject_fac, bs = 're') +
              s(age, subject_fac, bs = 're'), 
            data = dat_pred_caud2, 
            na.action = na.exclude, 
            method = "fREML")

summary(test2)
draw(test2)
plot(ggpredict(test2, terms = c("age")))
plot(ggpredict(test2, terms = c("drinking_group_ord")))
plot(ggpredict(test2, terms = c("age", "drinking_group_ord")))

#Preds
dat_pred_caud2$predicted <-  predict(test2, newdata = dat_pred_caud2, type = "response")




#If I want to color lines accordingly, I need to create average age by drinking-phase.
cutoff_age <- dat_pred_caud2 %>%
  filter(drinking_phase == "After") %>%
  group_by(subject) %>%
  summarize(min_after_age = min(age), .groups = "drop") %>%
  summarize(avg_cutoff = mean(min_after_age, na.rm = TRUE)) %>%
  pull(avg_cutoff)

dat_pred_caud2 <- dat_pred_caud2 %>%
  mutate(phase_age = case_when(
    drinking_phase == "Never" ~ "Never",
    age < cutoff_age ~ "Before",
    age >= cutoff_age ~ "After"
  ))

caud2 <- ggplot(dat_pred_caud2, aes(x = age, y = predicted, group = subject, color = drinking_phase)) +
  #geom_point(alpha = 0.2, size = 0.05) +
  geom_line(alpha = 0.2, size = 0.1) +
      scale_color_manual(values = c(
    "After" = "#7500b3ff",
    "Before" = "black", 
    "Never" = "black"))   +   
    new_scale_color() +
  new_scale_fill() +
  geom_smooth(aes(group = phase_age, color = phase_age), size = 0.3, show.legend = FALSE) +
  labs(y = "", x = "") +
    scale_color_manual(values = c(
    "After" = "#7500b3ff",  
    "Before" = "black", 
    "Never" = "black"
  )) +
    facet_wrap(~drinking_group, nrow = 1) + 
 # geom_vline(xintercept = cutoff_age, linetype = "dashed", size = 0.2) +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    legend.position = "none",
    strip.text = element_blank(),
    panel.background = element_blank(),   # remove gray panel
    panel.grid = element_blank(),         # remove grid lines
    plot.background = element_blank()     # remove outer gray
  )

caud2

ggsave(filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_binge/caud_pred_ban_plot.pdf", 
      plot = caud2,
       device = "pdf", 
       dpi = 300, 
      units = "mm",
       width = 90, 
       height = 60)
```


### Figure showing growth model colored by drinking onset (putamen)


```{r}
# Add predicted values to your data
dat_pred_put1 <- dat_long_ba %>%
  filter(seed_cluster == "putamen_cluster1") 

#Seeing if additional grouping variable works
dat_pred_put1 <- dat_pred_put1 %>%
  mutate(drinking_group = ifelse(drinking_phase == "Never", "Never", "Initiated"),
         drinking_group = factor(drinking_group, levels = c("Initiated", "Never")))

#Oredered?
dat_pred_put1$drinking_group_ord <- factor(dat_pred_put1$drinking_group, 
                             levels = c("Never", "Initiated"), 
                             ordered = TRUE)

#Seeing if there are actual trajectory differences
test3 <- bam(mean_value ~ 1 + sex_fac  +  drinking_group_ord +
              s(age, k = 5) +
              s(age, by = drinking_group_ord, k = 5) +
              s(subject_fac, bs = 're') +
              s(age, subject_fac, bs = 're'), 
            data = dat_pred_put1, 
            na.action = na.exclude, 
            method = "fREML")

summary(test3)
gam.vcomp(test3)
variance_comp(test3)
draw(test3)
plot(ggpredict(test3, terms = c("age")))
plot(ggpredict(test3, terms = c("drinking_group_ord")))
plot(ggpredict(test3, terms = c("age", "drinking_group_ord")))

#Preds
dat_pred_put1$predicted <-  predict(test3, newdata = dat_pred_put1, type = "response")




#If I want to color lines accordingly, I need to create average age by drinking-phase.
cutoff_age <- dat_pred_put1 %>%
  filter(drinking_phase == "After") %>%
  group_by(subject) %>%
  summarize(min_after_age = min(age), .groups = "drop") %>%
  summarize(avg_cutoff = mean(min_after_age, na.rm = TRUE)) %>%
  pull(avg_cutoff)

dat_pred_put1 <- dat_pred_put1 %>%
  mutate(phase_age = case_when(
    drinking_phase == "Never" ~ "Never",
    age < cutoff_age ~ "Before",
    age >= cutoff_age ~ "After"
  ))

put1 <- ggplot(dat_pred_put1, aes(x = age, y = predicted, group = subject, color = drinking_phase)) +
  #geom_point(alpha = 0.2, size = 0.05) +
  geom_line(alpha = 0.2, size = 0.1) +
      scale_color_manual(values = c(
    "After" = "#1b9e77",
    "Before" = "black", 
    "Never" = "black"))   +   
    new_scale_color() +
  new_scale_fill() +
  geom_smooth(aes(group = phase_age, color = phase_age), size = 0.3, show.legend = FALSE) +
  labs(y = "", x = "") +
    scale_color_manual(values = c(
    "After" = "#1b9e77",  
    "Before" = "black", 
    "Never" = "black"
  )) +
    facet_wrap(~drinking_group, nrow = 1) + 
 # geom_vline(xintercept = cutoff_age, linetype = "dashed", size = 0.2) +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    legend.position = "none",
    strip.text = element_blank(),
    panel.background = element_blank(),   # remove gray panel
    panel.grid = element_blank(),         # remove grid lines
    plot.background = element_blank()     # remove outer gray
  )

put1

ggsave(filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_binge/put_pred_ban_plot.pdf", 
      plot = put1,
       device = "pdf", 
       dpi = 300, 
      units = "mm",
       width = 90, 
       height = 60)
```



# Supplementary figures depicting difference smooths

I think I want to show the results from draw() and ggpredict() of the smooths next to each other, so [2,6] dimensions. This will be for supplement so no need to make incredibly fancy.

```{r}
#
#All of this would need to be ran if someone actually care about ages where smooths are different from each other.
#
#test4 <- bam(mean_value ~ 1 + sex_fac + drinking_group +
#              #s(age, k = 5) +
#              s(age, by = drinking_group, k = 5) +
#              s(subject_fac, bs = 're') +
#              s(age, subject_fac, bs = 're'), 
#            data = dat_pred_nacc1, 
#            na.action = na.exclude, 
#            method = "fREML")
#
#summary(test4)
##Function to calculate difference smooths
#test_diff4 <- difference_smooths(test4, select = c("s(age)"))
#
#tail(test_diff4)
#
#draw(test_diff4)

test_diff4

#NAcc
t1a <- plot(ggpredict(test, terms = c("age", "drinking_group_ord")), show_ci = FALSE) + 
  theme_classic() +
  labs(title = NULL,
       x = NULL,
       y = NULL) +
  theme(legend.position = "none")
t1a


t1b <- draw(test, select = "s(age):drinking_group_ordInitiated") + 
  theme_classic() +
  labs(title = NULL,
       x = NULL,
       y = NULL) +
  theme(legend.title = element_blank())
t1b

ggsave(filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_difference_smooths/nacc_smooths.pdf", 
      plot = t1a,
       device = "pdf", 
       dpi = 300, 
      units = "mm",
       width = 90, 
       height = 60)

ggsave(filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_difference_smooths/nacc_dif_smooths.pdf", 
      plot = t1b,
       device = "pdf", 
       dpi = 300, 
      units = "mm",
       width = 90, 
       height = 60)


#caud
t2a <- plot(ggpredict(test2, terms = c("age", "drinking_group_ord")), show_ci = FALSE) + 
  theme_classic() +
  labs(title = NULL,
       x = NULL,
       y = NULL) +
  theme(legend.position = "none")
t2a


t2b <- draw(test2, select = "s(age):drinking_group_ordInitiated") + 
  theme_classic() +
  labs(title = NULL,
       x = NULL,
       y = NULL) +
  theme(legend.title = element_blank())
t2b

ggsave(filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_difference_smooths/caudate_smooths.pdf", 
      plot = t2a,
       device = "pdf", 
       dpi = 300, 
      units = "mm",
       width = 90, 
       height = 60)

ggsave(filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_difference_smooths/caudate_dif_smooths.pdf", 
      plot = t2b,
       device = "pdf", 
       dpi = 300, 
      units = "mm",
       width = 90, 
       height = 60)


#put
t3a <- plot(ggpredict(test3, terms = c("age", "drinking_group_ord")), show_ci = FALSE) + 
  theme_classic() +
  labs(title = NULL,
       x = NULL,
       y = NULL) +
  theme(legend.position = "none")
t3a


t3b <- draw(test3, select = "s(age):drinking_group_ordInitiated") + 
  theme_classic() +
  labs(title = NULL,
       x = NULL,
       y = NULL) +
  theme(legend.title = element_blank())
t3b

ggsave(filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_difference_smooths/putamen_smooths.pdf", 
      plot = t3a,
       device = "pdf", 
       dpi = 300, 
      units = "mm",
       width = 90, 
       height = 60)

ggsave(filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_difference_smooths/putamen_dif_smooths.pdf", 
      plot = t3b,
       device = "pdf", 
       dpi = 300, 
      units = "mm",
       width = 90, 
       height = 60)
```


# Figure 2 work

This chunk of code can probably be moved to another script. But I want to create a figure that shows the brain growth models after the K-means clustering and averaging at individual level step. I'm thinking it might be useful to plot based on seeded region, and overlay the gam smooths. Probably makes the most sense to do this from the long data and subset based on seed.

```{r}
#Seeing if long makes more sense.
nacc_raw_plot <- ggplot(data = dat_long %>% 
                          filter(str_detect(seed_cluster, "nacc")), 
                        aes(x = age, 
                            y = mean_value,
                            colour = seed_cluster)) +
  geom_point(size = .05, alpha = .2, color = "grey60") + 
  #geom_line() +
  geom_smooth(aes(group = seed_cluster, 
                  color = seed_cluster),
              method ="gam",
              linewidth = 0.4) + 
  scale_color_manual(values = c("nacc_cluster1" = "#d95f02",
                                "nacc_cluster2" = "#e6ab02")) +
  labs(x = "",
       y = "") +
  theme_modern() +
  theme(
     legend.position = "none",
     plot.background = element_blank(),
     panel.background = element_blank(),
     axis.line = element_line(size = .2),
     axis.ticks = element_line(size = .2),
     #axis.line.x = element_line(color = "black"),
     #axis.line.y = element_line(color = "black"), 
     panel.grid = element_blank(),
     text = element_text(size = 5),
      axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5)
    ) 
nacc_raw_plot
#ggsave(filename = 
#         "H:/Dan/ncanda/R/figures/manuscript_figs/figure_2/nacc_raw_plot.pdf", 
#      plot = nacc_raw_plot,
#      device = "pdf", 
#      dpi = 500, 
#      units = "mm",
#      width = 50, 
#      height = 50)


#Seeing if long makes more sense.
caudate_raw_plot <- ggplot(data = dat_long %>% 
                          filter(str_detect(seed_cluster, "caudate")), 
                        aes(x = age, 
                            y = mean_value,
                            colour = seed_cluster)) +
  geom_point(size = .05, alpha = .2, color = "grey60") + 
  #geom_line() +
  geom_smooth(aes(group = seed_cluster, 
                  color = seed_cluster),
              method ="gam",
              linewidth = 0.4) + 
  scale_color_manual(values = c("caudate_cluster1" = "#7570b3",
                                "caudate_cluster2" = "#7500b3ff")) +
  labs(x = "",
       y = "") +
  theme_modern() +
  theme(
     legend.position = "none",
     plot.background = element_blank(),
     panel.background = element_blank(),
     axis.line = element_line(size = .2),
     axis.ticks = element_line(size = .2),
     #axis.line.x = element_line(color = "black"),
     #axis.line.y = element_line(color = "black"), 
     panel.grid = element_blank(),
     text = element_text(size = 5),
      axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5)
    ) 
caudate_raw_plot
ggsave(filename = 
         "H:/Dan/ncanda/R/figures/manuscript_figs/figure_2/caudate_raw_plot.pdf", 
      plot = caudate_raw_plot,
      device = "pdf", 
      dpi = 500, 
      units = "mm",
      width = 50, 
      height = 50)

#Seeing if long makes more sense.
putamen_raw_plot <- ggplot(data = dat_long %>% 
                          filter(str_detect(seed_cluster, "putamen")), 
                        aes(x = age, 
                            y = mean_value,
                            colour = seed_cluster)) +
  geom_point(size = .05, alpha = .2, color = "grey60") + 
  #geom_line() +
  geom_smooth(aes(group = seed_cluster, 
                  color = seed_cluster),
              method ="gam",
              linewidth = 0.4) + 
  scale_color_manual(values = c("putamen_cluster1" = "#1b9e77",
                                "putamen_cluster2" = "#1b3877ff")) +
  labs(x = "",
       y = "") +
  theme_modern() +
  theme(
     legend.position = "none",
     plot.background = element_blank(),
     panel.background = element_blank(),
     axis.line = element_line(size = .2),
     axis.ticks = element_line(size = .2),
     #axis.line.x = element_line(color = "black"),
     #axis.line.y = element_line(color = "black"), 
     panel.grid = element_blank(),
     text = element_text(size = 5),
      axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5)
    ) 
putamen_raw_plot
ggsave(filename = 
         "H:/Dan/ncanda/R/figures/manuscript_figs/figure_2/putamen_raw_plot.pdf", 
      plot = putamen_raw_plot,
      device = "pdf", 
      dpi = 500, 
      units = "mm",
      width = 50, 
      height = 50)
#50/50 seems right.
```


Dr. Parr would like regions of significant from these models, in particular, the NAcc cluster 1 results.

```{r}
#Explicitly setting factors
dat$subject_fac <- as.factor(dat$subject_fac)
dat$sex_fac <- as.factor(dat$sex_fac)

parr <- bam(nacc_cluster1 ~ 1 + sex_fac + s(age, k = 5) + s(subject_fac, bs = "re") + s(age, subject_fac, bs = 're'), data = dat)
summary(parr)

plot(ggpredict(parr, terms = "age"))

#Extract derivatives
d <- derivatives(parr,
                 interval = "simultaneous",
                 n_sim = 10000,
                 n = 200)
#Plot derivatives
draw(d)

#Calculate regions of significance, where CI != 0
d <- d %>% mutate(sig = !(0>.lower_ci & 0<.upper_ci))

#Derivative is !0
#No change from 12 - 	14.64201
#Change from 14.72874 - 	20.19296
#No change from 20.27970 - 23.14191
#Change from 23.22864 - 24.96332
#No change from 25.05005 onwards.

#Keep in mind this data includes both drinkers and non-drinkers, so tail at older ages is probably being warped a bit.

#23.22864

```