---
title: "ncanda_cleaning"
author: "Daniel Petrie"
date: "2024-08-20"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r Global}
#Packages
library(tidyverse) #Wrangling
library(dplyr) #Wrangling
library(stringr) #For str_replace
library(ggplot2)
library(psych)
library(see) #theme_modern()
#WD. Change whether through H: or on different sever
setwd("H:/Dan/ncanda/R")

#CDDR data. I think I want to use the "youth report" version for now because it has ages.
#Baseline through year 8
cddr0_7 <- read.csv("H:/Raw/Datasets/NCANDA/followup_7y/NCANDA_RELEASE_7Y_REDCAP_MEASUREMENTS_V01/summaries/redcap/youthreport1_cddr.csv", header = TRUE)
#Year 9
cddr8 <- read.csv("H:/Raw/Datasets/NCANDA/followup_8y/NCANDA_RELEASE_8Y_REDCAP_MEASUREMENTS_V01/summaries/redcap/youthreport1_cddr.csv", header = TRUE)

#Dataset to extract sex
dem0_7 <- read.csv("H:/Raw/Datasets/NCANDA/followup_7y/NCANDA_RELEASE_7Y_REDCAP_MEASUREMENTS_V01/summaries/redcap/demographics.csv", header = TRUE)
```

Cleaning 
add youthreport1_cddr15

```{r CDDR Cleaining}
#Vector of alcohol variable names
subset_var <- c("subject", "visit", "youthreport1_age", "youthreport1_cddr15", "youthreport1_cddr16",
                "youthreport1_cddr17", "youthreport1_cddr18", "youthreport1_cddr19", "youthreport1_cddr21",
                "youthreport1_cddr22", "youthreport1_cddr23", "youthreport1_cddr26", "youthreport1_cddr27",
                "youthreport1_cddr30", "youthreport1_cddr31")

#Binding data, arrange by visit, select alcohol variables, rename.
cddr0_8 <- bind_rows(cddr0_7, cddr8) %>% 
  arrange(subject, visit) %>%
  dplyr::select(all_of(subset_var)) %>%
  rename_with(~ str_replace(., "^youthreport1_", ""), starts_with("youthreport1_"))

# List of columns to modify
variable_list <- c("cddr15", "cddr16", "cddr17", "cddr18",
                   "cddr19", "cddr21", "cddr22", "cddr23",
                   "cddr26", "cddr27", "cddr30", "cddr31")

#Recoding some NAs to 0's to reflect an absence of drinking behavior and not a true missing value.
cddr0_8 <- cddr0_8 %>%
  mutate(across(all_of(variable_list), ~ ifelse(is.na(.),
                                                ifelse(is.na(age), NA, 0),
                                                .)))
#Subset demographic data for sex.
dem0_7_sub <- dem0_7 %>%
  dplyr::select(subject, visit, sex)
#Adding sex
cddr0_8 <- merge(cddr0_8, dem0_7_sub, by = c("subject", "visit"))
```


**CDDR variable descriptions**
11 variables total.

THIS DOES NOT SEEM TO BE ADDED! cddr15: How old were you when you first drank ALCOHOL? (beer, wine, or hard liqour; more than a taste/sip) (code age to nearest quarter year using decimals, e.g. 141/4 = 14.25)

cddr16: How old were you when you first drank ALCOHOL regularly? (at least once a week; beer, wine, or hard liquor)  

cddr17: When was the last time you drank ALCOHOL? (0=today, 1=yesterday, etc.)

cddr18: How many ALCOHOLIC DRINKS did you drink on that occasion?

cddr19: In your lifetime, how many days did you drink ALCOHOL (beer, wine, hard liquor)?

cddr21: During the past year, (so out of 365 days), how many days did you drink ALCOHOL (beer, wine, hard liquor)? (0-365)

cddr22: During the last 30 days, how many days did you drink ALCOHOL? (0-30)

cddr23: Over the last month, in the average 24-hour period you were drinking, how many ALCOHOLIC DRINKS did you have?

cddr26: During the past year (or 365 days), in the average 24-hour period you were drinking how many total drinks did you have? 

cddr27: Have you ever consumed 4+ (females) / 5+ (males) drinks within an occasion? (y/n)

cddr30: During the past year, how many times have you consumed 4+ (females) / 5+ (males) drinks within an occasion? (0-365) 

cddr31: During the last 30 days, how many times have you consumed 4+ (females) / 5+ (males) drinks within an occasion? (0-30)


```{r}
# Categorize drinking levels using month values
cddr0_8_cat <- cddr0_8 %>%
  mutate(drinking_category_month = case_when(
    (sex == "M" & cddr23 <= 2 & cddr31 == 0) |
      (sex == "F" & cddr23 <= 1 & cddr31 == 0) ~ "no_low",
    (sex == "M" & cddr23 > 2 & cddr23 <= 4 | cddr31 < 5) |
      (sex == "F" & cddr23 > 1 & cddr23 <= 3 | cddr31 < 5) ~ "moderate",
    (sex == "M" & (cddr23 > 4 | cddr31 >= 5)) |
      (sex == "F" & (cddr23 > 3 | cddr31 >= 5)) ~ "heavy",
    TRUE ~ "unclassified" # Catch any unexpected cases
  )) %>% 
  filter(drinking_category_month != "unclassified") %>% 
  mutate(
    drinking_category_month = factor(drinking_category_month,
                                     levels = c("no_low", "moderate", "heavy")))
```

Write csv in clean_data

```{r}
#write.csv(cddr0_8, file = "./clean_data/cddr0_8.csv", row.names = FALSE)
```