---
title: "ncanda_merging"
author: "Daniel Petrie"
date: "2024-10-17"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library("ggplot2") #For plotting
library("GGally") #ggpairs()
library("tidyverse") #Wranglin
library("dplyr") #Wranglin
library("lme4")
library(lmerTest)
#WD
setwd("H:/Dan/ncanda/R")

#Data
#Demographics
demographics <- read.csv("clean_data/demographics.csv", header = TRUE)

#Drinking measures
cddr0_8 <- read.csv("clean_data/cddr0_8.csv", header = TRUE)

#OCD measures
ysr_asr0_8 <- read.csv("clean_data/ysr_asr0_8.csv", header = TRUE)

#Tissue iron
tat2_harox <- read.csv("clean_data/tat2_harox.csv", header = TRUE)

#Functional Connectivity measures (these will change)
#roi_roi_aal3 <- read.csv("clean_data/roi_roi_aal3.csv", header = TRUE)
putamen_seed <- read.csv("clean_data/putamen_seed_clean.csv")
caudate_seed <- read.csv("clean_data/caudate_seed_clean.csv")
nacc_seed <- read.csv("clean_data/nacc_seed_clean.csv")
seed <- read.csv("clean_data/gordon_network_seed_clean.csv", header = TRUE)

#These probably won't
high_motion_subjects <- read.csv("clean_data/high_motion_subjects.csv", header = TRUE)
```


This script creates four different non-preprocesseed data frames, and can be adopted to clean most types of functional connectivity data.
1.) putamen seed data
2.) caudate seed data
3.) nacc seed data
4.) network-level data across all seeded regions.

Data that I want from each data set:
1) demographics: c("subject", "visit", "sex", "visit_age", "site", *"scanner"*)

2) cddr0_8: c("subject", "visit" "cddr16", "cddr17", "cddr18", "cddr19", 
"cddr21", "cddr22", "cddr23", "cddr26", "cddr27", "cddr30", "cddr31")

3) ysr_asr0_8: c("subject", "visit", "ocd_raw", "ocd_t", "ocd_pct", "aseba")

4) roi_roi_aal3: c("subject", "visit", "put_precg", "put_poscg", "put_sma", "put_caud", 
"put_ofc", "put_vmpfc", "put_dlpfc", "precg_poscg", "precg_sma", 
"precg_caud", "precg_ofc", "precg_vmpfc", "precg_dlpfc", "poscg_sma", 
"poscg_caud", "poscg_ofc", "poscg_vmpfc", "poscg_dlpfc", "sma_caud", 
"sma_ofc", "sma_vmpfc", "sma_dlpfc", "caud_ofc", "caud_vmpfc", 
"caud_dlpfc", "ofc_vmpfc", "ofc_dlpfc", "vmpfc_dlpfc")

5) high_motion_subjects (to remove at the end): c("id", "count", "scanner", "age", "visit", "remove") 


Subsetting and renaming variables for all datasets. I'm merging on visitnum_numeric and id.

```{r}
#demographics
demographics_sub <- demographics %>% 
  rename(age = visit_age)

cddr0_8_sub <- cddr0_8 %>% 
  select("subject", "visit", "cddr16", "cddr17", 
         "cddr18", "cddr19", "cddr21", "cddr22", 
         "cddr23", "cddr26", "cddr27", "cddr30", 
         "cddr31", "drinking_category_month")

ysr_asr0_8_sub <- ysr_asr0_8 %>% 
  select("subject", "visit", "ocd_raw", 
         "ocd_t", "ocd_pct", "aseba")

tat2_harox_sub <- tat2_harox %>%
  select(-visit_numeric) %>%
  rename(subject = id) %>%
  mutate(subject = paste0("NCANDA_", subject)) %>%
  rename_with(~ paste0(., "_iron"), .cols = -(1:2))

#This should work across neuroimaging measures
#ROI-ROI
#roi_roi_aal3_sub <- roi_roi_aal3 %>% 
#  mutate(subject = paste0("NCANDA_", subject)) %>%
#  rename(visit_numeric = visit) %>%
#  mutate(visit = case_when(
#    visit_numeric == 0 ~ "baseline",
#    visit_numeric == 1 ~ "followup_1y",
#    visit_numeric == 2 ~ "followup_2y",
#    visit_numeric == 3 ~ "followup_3y",
#    visit_numeric == 4 ~ "followup_4y",
#    visit_numeric == 5 ~ "followup_5y",
#    visit_numeric == 6 ~ "followup_6y",
#    visit_numeric == 7 ~ "followup_7y",
#    visit_numeric == 8 ~ "followup_8y",
#    TRUE ~ NA_character_  # Handle any unexpected visit values
#  )) %>%
#  select(-visit_numeric)

#Only selecting network level columns for now.
#Seeded at Gordon ROI level
putamen_seed_sub <- putamen_seed %>% 
  mutate(subject = paste0("NCANDA_", subject)) %>%
  #rename(visit_numeric = visit) %>%
  mutate(visit = case_when(
    visit_numeric == 0 ~ "baseline",
    visit_numeric == 1 ~ "followup_1y",
    visit_numeric == 2 ~ "followup_2y",
    visit_numeric == 3 ~ "followup_3y",
    visit_numeric == 4 ~ "followup_4y",
    visit_numeric == 5 ~ "followup_5y",
    visit_numeric == 6 ~ "followup_6y",
    visit_numeric == 7 ~ "followup_7y",
    visit_numeric == 8 ~ "followup_8y",
    TRUE ~ NA_character_  # Handle any unexpected visit values
  )) %>%
  select(-visit_numeric)

caudate_seed_sub <- caudate_seed %>% 
  mutate(subject = paste0("NCANDA_", subject)) %>%
  #rename(visit_numeric = visit) %>%
  mutate(visit = case_when(
    visit_numeric == 0 ~ "baseline",
    visit_numeric == 1 ~ "followup_1y",
    visit_numeric == 2 ~ "followup_2y",
    visit_numeric == 3 ~ "followup_3y",
    visit_numeric == 4 ~ "followup_4y",
    visit_numeric == 5 ~ "followup_5y",
    visit_numeric == 6 ~ "followup_6y",
    visit_numeric == 7 ~ "followup_7y",
    visit_numeric == 8 ~ "followup_8y",
    TRUE ~ NA_character_  # Handle any unexpected visit values
  )) %>%
  select(-visit_numeric)

nacc_seed_sub <- nacc_seed %>% 
  mutate(subject = paste0("NCANDA_", subject)) %>%
  #rename(visit_numeric = visit) %>%
  mutate(visit = case_when(
    visit_numeric == 0 ~ "baseline",
    visit_numeric == 1 ~ "followup_1y",
    visit_numeric == 2 ~ "followup_2y",
    visit_numeric == 3 ~ "followup_3y",
    visit_numeric == 4 ~ "followup_4y",
    visit_numeric == 5 ~ "followup_5y",
    visit_numeric == 6 ~ "followup_6y",
    visit_numeric == 7 ~ "followup_7y",
    visit_numeric == 8 ~ "followup_8y",
    TRUE ~ NA_character_  # Handle any unexpected visit values
  )) %>%
  select(-visit_numeric)

#Network level seeded connectivity
seed_sub <- seed %>% 
  mutate(subject = paste0("NCANDA_", subject)) %>%
  #rename(visit_numeric = visit) %>%
  mutate(visit = case_when(
    visit_numeric == 0 ~ "baseline",
    visit_numeric == 1 ~ "followup_1y",
    visit_numeric == 2 ~ "followup_2y",
    visit_numeric == 3 ~ "followup_3y",
    visit_numeric == 4 ~ "followup_4y",
    visit_numeric == 5 ~ "followup_5y",
    visit_numeric == 6 ~ "followup_6y",
    visit_numeric == 7 ~ "followup_7y",
    visit_numeric == 8 ~ "followup_8y",
    TRUE ~ NA_character_
  )) %>%
  select(-visit_numeric)



high_motion_subjects_sub <- high_motion_subjects %>% 
  mutate(subject = gsub("_.*", "", id)) %>% 
  mutate(subject = paste0("NCANDA_", subject)) %>%
  select(subject, count, scanner, visit, remove)
```


I want the neuroimaging data to have the least amount of missingness.


```{r Merging}
#putamen
#Keeping all imaging data
temp0_putamen <- left_join(tat2_harox_sub, putamen_seed_sub, by = c("subject", "visit"))
temp1_putamen <- left_join(temp0_putamen, demographics_sub, by = c("subject", "visit"))
temp2_putamen <- left_join(temp1_putamen, cddr0_8_sub, by = c("subject", "visit"))
temp3_putamen <- left_join(temp2_putamen, ysr_asr0_8_sub, by = c("subject", "visit"))

#Rearrange for ease of viewing
temp4_putamen <- temp3_putamen %>%
  select(subject, visit, age, sex, site, scanner, everything())

#caudate
#Keeping all imaging data
temp0_caudate <- left_join(tat2_harox_sub, caudate_seed_sub, by = c("subject", "visit"))
temp1_caudate <- left_join(temp0_caudate, demographics_sub, by = c("subject", "visit"))
temp2_caudate <- left_join(temp1_caudate, cddr0_8_sub, by = c("subject", "visit"))
temp3_caudate <- left_join(temp2_caudate, ysr_asr0_8_sub, by = c("subject", "visit"))

#Rearrange for ease of viewing
temp4_caudate <- temp3_caudate %>%
  select(subject, visit, age, sex, site, scanner, everything())

#nacc
#Keeping all imaging data
temp0_nacc <- left_join(tat2_harox_sub, nacc_seed_sub, by = c("subject", "visit"))
temp1_nacc <- left_join(temp0_nacc, demographics_sub, by = c("subject", "visit"))
temp2_nacc <- left_join(temp1_nacc, cddr0_8_sub, by = c("subject", "visit"))
temp3_nacc <- left_join(temp2_nacc, ysr_asr0_8_sub, by = c("subject", "visit"))

#Rearrange for ease of viewing
temp4_nacc <- temp3_nacc %>%
  select(subject, visit, age, sex, site, scanner, everything())

#Network
#Keeping all imaging data
temp0 <- left_join(tat2_harox_sub, seed_sub, by = c("subject", "visit"))
temp1 <- left_join(temp0, demographics_sub, by = c("subject", "visit"))
temp2 <- left_join(temp1, cddr0_8_sub, by = c("subject", "visit"))
temp3 <- left_join(temp2, ysr_asr0_8_sub, by = c("subject", "visit"))

#Rearrange for ease of viewing
temp4 <- temp3 %>%
  select(subject, visit, age, sex, site, scanner, everything())
```

Removing high motion subjects

```{r Remove High Motion}
temp5_putamen <- anti_join(temp4_putamen, 
                   high_motion_subjects_sub, 
                   by = c("subject", "visit"))
temp5_caudate <- anti_join(temp4_caudate, 
                   high_motion_subjects_sub, 
                   by = c("subject", "visit"))
temp5_nacc <- anti_join(temp4_nacc, 
                   high_motion_subjects_sub, 
                   by = c("subject", "visit"))
temp5 <- anti_join(temp4, 
                   high_motion_subjects_sub, 
                   by = c("subject", "visit"))
```



Write csv. Note that pre-processing will happen in a different script.

```{r}
#roi-roi
#write.csv(temp5, file = "clean_data/no_preprocess_full.csv", row.names = FALSE)

#seed Gordon ROI
write.csv(temp5_putamen, file = "clean_data/no_preprocess_putamen_seed.csv", row.names = FALSE)
write.csv(temp5_caudate, file = "clean_data/no_preprocess_caudate_seed.csv", row.names = FALSE)
write.csv(temp5_nacc, file = "clean_data/no_preprocess_nacc_seed.csv", row.names = FALSE)

#seed Gordon network
write.csv(temp5, file = "clean_data/no_preprocess_gordon_network_seed.csv", row.names = FALSE)
```

