---
title: "test_drinking_categories"
author: "Daniel Petrie"
date: "2025-02-28"
output: html_document
editor_options: 
  chunk_output_type: console
---

Global

```{r}
library(lcmm)
library(dplyr)
library(cluster)
library(factoextra)
library(GGally)
library(NbClust)
#Clean seed data
dat_putamen <- read.csv("./clean_data/full_putamen_seed_020425.csv", header = TRUE)
```


Cleaning chunk

```{r}
dat_putamen_sub <- dat_putamen %>%
  dplyr::select(subject, age,	sex,	
                visit,	visit_numeric, 
                matches("^cddr\\d{2}$"), drinking_category_month) %>%
  filter(!(subject == "NCANDA_S00991" & visit_numeric == "2"))

```

I want to try and create person-level drinking categories that do not vary over time.

cddr21: During the past year, (so out of 365 days), how many days did you drink ALCOHOL (beer, wine, hard liquor)? (0-365)

cddr26: During the past year (or 365 days), in the average 24-hour period you were drinking how many total drinks did you have? 

cddr30: During the past year, how many times have you consumed 4+ (females) / 5+ (males) drinks within an occasion? (0-365)

K-means clustering

Focusing on summary statistics that capture different aspects of drinking patterns without requiring explicit modeling of trajectories. These features (averages, maximums, volatility, and proportions) will likely capture meaningful differences between drinking groups without making assumptions about the shape of trajectories.

NOTE: When fussing with feature selection, there are some instances were there was very little variability. This is bad because k-means clustering relies on differences between data points to form meaningful groups. K-means are based on Euclidean distance, so a feature with little variability does not influence distances between points.

NOTE on acceleration: Captures escalation patterns: If drinking accelerates quickly in adolescence, it may signal higher risk for problematic substance use. Someone with a steady drinking increase will have low accerleration, while someone with irregular spikes will have high acceleration.

NOTE: I am going to stick with configuation below where I just use days and binge episodes.

```{r}
# Create feature matrix
subject_features <- dat_putamen_sub %>%
  group_by(subject) %>%
  summarize(
  # Average drinking behaviors (frequency)
    avg_days = mean(cddr21, na.rm = TRUE),
    #avg_amount = mean(cddr26, na.rm = TRUE),  #These seem to not be differentiated enough
    avg_binge = mean(cddr30, na.rm = TRUE),
  #Max drinking behaviors (magnitude)
    max_days = max(cddr21, na.rm = TRUE),
    #max_amount = max(cddr26, na.rm = TRUE),  #These seem to not be differentiated enough
    max_binge = max(cddr30, na.rm = TRUE),
  # Standard deviation captures volatility in drinking
    volatility_days = ifelse(length(na.omit(cddr21)) > 1, sd(cddr21, na.rm = TRUE), 0),
    #volatility_amount = ifelse(length(na.omit(cddr26)) > 1, sd(cddr26, na.rm = TRUE), 0),
    volatility_binge = ifelse(length(na.omit(cddr30)) > 1, sd(cddr30, na.rm = TRUE), 0),
  # Area under curve captures total drinking exposure
    auc_days = sum(cddr21, na.rm = TRUE),
    #auc_amount = sum(cddr26, na.rm = TRUE),
    auc_binge = sum(cddr30, na.rm = TRUE)#,
  #Trajectory (i.e., age) features
  # Age of onset (first non-zero drinking)
    #onset_age = ifelse(any(cddr21 > 0, na.rm = TRUE), min(age[cddr21 > 0], na.rm = TRUE), 0),
  ## Age at peak consumption
    #peak_age = ifelse(any(cddr21 > 0, na.rm = TRUE), age[which.max(cddr21)], 0),
  ## Rate of acceleration (change in slope)
    #acceleration = ifelse(length(unique(cddr21[!is.na(cddr21)])) > 1, 
    #                      sd(diff(cddr21)/diff(age), na.rm = TRUE), 0)
  )

describe(subject_features[,-1])

#Checking NAs
colSums(is.na(subject_features))
#50 or 53 subjects only have one time point. Removing them for classification



#Omit instances where
#subject_features <- na.omit(subject_features)


# Standardize features
features_scaled <- scale(subject_features[,-1])

describe(features_scaled[,-1])
```


Trying to determine amount of clusters to start at.

```{r}
# Elbow method
fviz_nbclust(features_scaled, kmeans, method = "wss") +
  labs(title = "Elbow Method for Optimal k")

#Using NbClust package
# Compare several k-means models with different metrics
# NbClust provides 30 indices for determining number of clusters
nb <- NbClust(features_scaled, distance = "euclidean",
              min.nc = 2, max.nc = 10, method = "kmeans")

# Histogram of optimal number of clusters
barplot(table(nb$Best.n[1,]),
        xlab = "Number of Clusters",
        ylab = "Number of Criteria",
        main = "Number of Clusters Chosen by 30 Criteria")
```


K-Means Clustering Chunk

```{r K-Means}
#Set seed for reproducable reuslts
set.seed(1738) #Fetty Wap

# K-means clustering with 3 groups
km_result <- kmeans(features_scaled, centers = 3)

#seing table of results
table(km_result$cluster)
# Add cluster assignments back to data

subject_features$cluster <- km_result$cluster
```


Viz for cluster assignments

```{r}
# Basic cluster visualization on first two PCs
fviz_cluster(list(data = features_scaled, cluster = km_result$cluster),
             ellipse.type = "convex",
             repel = TRUE,
             ggtheme = theme_modern(),
             palette = c("#002643", "#840032", "#E59500")) +
  labs(title = "Drinking Pattern Clusters")

#Other way of visualizing.
fviz_cluster(km_result, 
             data = features_scaled, 
             geom = "point", 
             ellipse.type = "convex",
             ggtheme = theme_modern(),
             palette = c("#002643", "#840032", "#E59500")) +
  labs(title = "Drinking Pattern Clusters")

# More detailed visualization with original features
# Parallel coordinates plot showing feature patterns by cluster
features_with_clusters <- cbind(as.data.frame(features_scaled), 
                               cluster = as.factor(km_result$cluster))

features_clusters <- ggparcoord(features_with_clusters, 
           columns = 1:ncol(features_scaled), 
           groupColumn = "cluster",
           scale = "uniminmax",
           alphaLines = 0.1) +
  theme_minimal() +
  labs(title = "Feature Patterns by Drinking Cluster",
       x = "Features",
       y = "Standardized Value") +
  scale_color_manual(values = c("#002643", "#840032", "#E59500")) +  
  theme(
    plot.title = element_text(size = 8),
    axis.title = element_text(size = 8),
    axis.text = element_text(size = 8),
    axis.line = element_line(linewidth = .2),
    axis.ticks = element_line(linewidth = .2),
    #legend.title = element_text(size = 8),
    #legend.text = element_text(size = 8),
    legend.position = "none",
    panel.background = element_blank()
  )
features_clusters
ggsave(plot = features_clusters,
       filename = "H:/Dan/ncanda/R/figures/features_clusters.pdf", 
       device = "pdf", 
       dpi = 500,
       units = "mm",
       width = 120, 
       height = 80)
```


Trajectories by cluster

```{r}
# After assigning clusters to original data
dat_putamen_sub$cluster <- subject_features$cluster[match(dat_putamen_sub$subject, subject_features$subject)]

dat_putamen_sub <- dat_putamen_sub %>%
  mutate(cluster_fac = factor(cluster, 
                                   levels = c(1, 3, 2),  # New order (example)
                                   labels = c("Low Use", "Moderate Use", "High Use")))  # New names

cluster_colors <- c("Low Use" = "#002643", "Moderate Use" = "#E59500", "High Use" =  "#840032")

# Plot individual trajectories colored by cluster
cddr21_clust <- ggplot(dat_putamen_sub %>% filter(cluster != "NA"), 
                       aes(x = age, 
                           y = cddr21, 
                           group = subject, 
                           color = as.factor(cluster_fac))) +
  geom_line(alpha = 0.2,
            linewidth = 0.2) +
  geom_smooth(aes(group = cluster_fac), method = "gam", se = TRUE, linewidth = 0.5) +
  facet_wrap(~ cluster_fac) +
  #theme_modern() +
  scale_color_manual(values = cluster_colors) +
  labs(title = "Drinking Days by Cluster",
       x = "Age (years)",
       y = "Drinking Days Past Year",
       color = "Cluster") +
    theme(
    plot.title = element_text(size = 6),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    strip.text = element_text(size = 6, face = "plain"),
    strip.background = element_blank(),
    #legend.title = element_text(size = 8),
    #legend.text = element_text(size = 8),
    legend.position = "none",
    panel.background = element_blank()
  )
cddr21_clust
ggsave(plot = cddr21_clust,
       filename = "H:/Dan/ncanda/R/figures/RSA_PittPsych_2025/cddr21_clust.pdf", 
       device = "pdf", 
       dpi = 500,
       units = "mm",
       width = 90, 
       height = 45)

cddr30_clust <- ggplot(dat_putamen_sub %>% filter(cluster != "NA"), 
                       aes(x = age, 
                           y = cddr30, 
                           group = subject, 
                           color = as.factor(cluster_fac))) +
  geom_line(alpha = 0.2,
            linewidth = 0.2) +
  geom_smooth(aes(group = cluster_fac), method = "gam", se = TRUE, linewidth = 0.5) +
  facet_wrap(~ cluster_fac) +
 # theme_modern() +
  scale_color_manual(values = cluster_colors) +
  labs(title = "Binge Days by Cluster",
       x = "Age (years)",
       y = "Binge Days Past Year",
       color = "Cluster") +
    theme(
    plot.title = element_text(size = 6),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    strip.text = element_text(size = 6, face = "plain"),
    strip.background = element_blank(),
    #legend.title = element_text(size = 8),
    #legend.text = element_text(size = 8),
    legend.position = "none",
    panel.background = element_blank()
  )
cddr30_clust
ggsave(plot = cddr30_clust,
       filename = "H:/Dan/ncanda/R/figures/RSA_PittPsych_2025/cddr30_clust.pdf", 
       device = "pdf", 
       dpi = 500,
       units = "mm",
       width = 90, 
       height = 45)



# Heatmap of drinking patterns
dat_putamen_sub %>%
  filter(cluster != "NA") %>%
  group_by(cluster, age = round(age)) %>%
  summarize(mean_days = mean(cddr21, na.rm = TRUE)) %>%
  ggplot(aes(x = age, y = as.factor(cluster), fill = mean_days)) +
  geom_tile() +
  scale_fill_viridis_c() +
  theme_minimal() +
  labs(title = "Drinking Intensity by Age and Cluster",
       x = "Age",
       y = "Cluster",
       fill = "Avg. Drinking Days")

# Improved heatmap with better handling of missing data
dat_putamen_sub %>%
  filter(!is.na(cluster)) %>%
  group_by(cluster, age = round(age)) %>%
  summarize(mean_days = mean(cddr21, na.rm = TRUE),
            n_obs = n()) %>%
  # Filter to ensure minimum number of observations per cell
  filter(n_obs >= 3) %>%  # Only include cells with at least 3 observations
  ggplot(aes(x = age, y = as.factor(cluster), fill = mean_days)) +
  geom_tile() +
  scale_fill_viridis_c(na.value = "white") +  # Explicitly set NA color
  theme_minimal() +
  labs(title = "Drinking Intensity by Age and Cluster",
       x = "Age",
       y = "Cluster",
       fill = "Avg. Drinking Days")


#Looking at age range for each cluster assignment
ggplot(dat_putamen_sub %>% filter(cluster == "2"), aes(x = age, y = cddr30, group = subject)) +
  geom_point(alpha = 0.5, color = "blue") +  # Scatter plot with transparency
  geom_line() +
  geom_smooth(method = "loess", se = TRUE, color = "red", aes(group = 1)) +  # Smoothed trend line
  #facet_wrap(~ cluster) +  # Facet by cluster
  theme_minimal() +
  labs(title = "Age vs. cddr21 by Cluster",
       x = "Age",
       y = "Drinking Days Past Year") +
  theme(strip.text = element_text(size = 10))  # Adjust facet label size

```

I think this should be fine. Saving cluster assignments as a .csv file. Note that I need to reclean in other scripts.

```{r}
#write.csv(subject_features, file = "./clean_data/clusters.csv", row.names = FALSE)
```

