---
title: "nicotine_cannabis_specificity"
output: html_document
date: "2025-12-10"
editor_options: 
  chunk_output_type: console
---

```{r Global}
#Packages
library(tidyverse) #Wrangling
library(dplyr) #Wrangling
library(stringr) #For str_replace
library(ggplot2)
library(psych)
library(see) #theme_modern()
library(mgcv)
library(ggeffects)
library(emmeans)
#WD. Change whether through H: or on different sever
setwd("H:/Dan/ncanda/R")

#Clean data
dat_long <- read.csv("./clean_data/fc_clusters_alcohol_long_101025.csv", header = TRUE)
dat <- read.csv("./clean_data/fc_clusters_alcohol_101025.csv", header = TRUE)

#Nicotine/cannabis data
nc_dat <- read.csv("./clean_data/cddr0_8_nicotine_cannabis.csv", header = TRUE)
```

Needs cleaned. I want below variables.

cddr3   In the past 30 days, how many days did you smoke cigarettes? (0-30)

cddr5   How many cigarettes have you smoked in the last year, so out of 365 days?

cddr56 In the last year ( so out of 365 days), how many days did you use MARIJUANA? (0-365)

cddr57 In the last 30 days, how many days did you use MARIJUANA? (0-30)

Sandy NCANDA comment: "habitual cannabis use (use 5+ days per week)"

```{r}
nc_dat_sub <- nc_dat %>%
  select(subject, visit, cddr5, cddr3, cddr56, cddr57)

dat_nc <- merge(dat, nc_dat_sub, by = c("subject", "visit"))

#dat_nc_cddr5 <- dat_nc %>%
#  filter(cddr5 < 366)

#I kinda want to see how many people endorse cannabis use 5+ days.
#21 days is about 5 per week
#q<-dat_nc %>% mutate(habcan = cddr57 > 21) 
#%>% group_by(habcan) %>% summarise(habcan = n())
#qtrue <- q %>%
#  filter(habcan == TRUE)
##115 people reported "habitual cannabis use.
#length(unique(qtrue$subject))
#
#test <- gam(cddr57 ~ 1 + s(age) + s(age, by = putamen_cluster1), data = qtrue)
#summary(test)
#plot(ggpredict(test, terms = c("age", "putamen_cluster1")))  
##remove above gam model/
#
#      "cddr3 ~ 1 + sex_fac + 
#       s(age, k = 5) + 
#       s(age, by = ", v, ", k = 5)"
#    )
#  )
```

Some plots getting a feel for distributions of these new variables. I think the nicotine month might be more useful.

```{r SU traj plots}
#Getting portion of people who answered 0
mean(dat_nc$cddr3 == 0)
mean(dat_nc$cddr5 == 0)
mean(dat_nc$cddr56 == 0)
mean(dat_nc$cddr57 == 0)

#
describe(dat_nc$cddr3)
summary(dat_nc$cddr5)
table(dat_nc$cddr3)
table(dat_nc$cddr5)


ggplot(dat_nc %>% filter(cddr5 < 10), aes(x = cddr5)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  labs(x = "Cigarettes smoked in past year (days)", y = "Number of subjects") +
  theme_minimal()

ggplot(dat_nc, aes(x = cddr3)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  labs(x = "Cigarettes smoked in past year (days)", y = "Number of subjects") +
  theme_minimal()

ggplot(dat_nc, aes(x = cddr57)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  labs(x = "Cannabis past month (days)", y = "Number of subjects") +
  theme_minimal()


#Violin by visit
ggplot(dat_nc %>% filter(cddr56 < 365), aes(x = visit, y = cddr56)) +
  geom_violin(fill = "lightgreen") +
  geom_jitter(width = 0.2, alpha = 0.5) +
  labs(y = "Cigarettes smoked in past year (days)", x = "Sex")

#Spag plots
ggplot(dat_nc %>% filter(cddr5 < 365), aes(x = age, 
                      y = cddr5, 
                      group = subject)) +
  geom_line(alpha = 0.2,
            linewidth = 0.2) +
  geom_smooth(aes(group = 1), method = "gam", se = TRUE, linewidth = 0.5)

ggplot(dat_nc, aes(x = age, 
                      y = cddr3, 
                      group = subject)) +
  geom_line(alpha = 0.2,
            linewidth = 0.2) +
  geom_smooth(aes(group = 1), method = "gam", se = TRUE, linewidth = 0.5)

ggplot(dat_nc, aes(x = age, 
                      y = cddr56, 
                      group = subject)) +
  geom_line(alpha = 0.2,
            linewidth = 0.2) +
  geom_smooth(aes(group = 1), method = "gam", se = TRUE, linewidth = 0.5)
```

# TVEM specificity

I need to run this for my cannabis and nicotine measures twice.

```{r Nicotine month TVEM}
by_vars <- c(
  "putamen_cluster1",
  "putamen_cluster2",
  "caudate_cluster1",
  "caudate_cluster2",
  "nacc_cluster1",
  "nacc_cluster2"
)

# Storage lists
fits_nic <- list()
model_warnings_nic <- list()

for (v in by_vars) {

  # build formula
  formula <- as.formula(
    paste0(
      "cddr3 ~ 1 + sex_fac + 
       s(age, k = 5) + 
       s(age, by = ", v, ", k = 5)"
    )
  )

  # fit model
  fit <- gamm(
    formula,
    random = list(subject = ~1, subject = ~0 + age),
    data = dat_nc,
    method = "REML",
    na.action = na.exclude
  )

  # store model
  fits_nic[[v]] <- fit

  # capture warnings produced during summary()
  w <- NULL
  invisible(
    withCallingHandlers(
      summary(fit$gam),
      warning = function(wrn) {
        w <<- c(w, conditionMessage(wrn))
        invokeRestart("muffleWarning")
      }
    )
  )

  model_warnings_nic[[v]] <- w
}

# Optional: print GAM summaries
lapply(fits_nic, function(x) summary(x$gam))
model_warnings_nic

#Explicity showing pvalues

# Extract only "by" smooth p-values across all models
by_smooth_pvals_nic <- map_df(seq_along(fits_nic), function(i) {
  s <- summary(fits_nic[[i]]$gam)
  
  # Keep only smooths where the term name contains "by="
  smooth_table <- as.data.frame(s$s.table) %>%
    mutate(term = rownames(s$s.table),
           model = i) %>%
    filter(grepl(":", term)) %>%
    select(model, term, edf = edf, F, p_value = `p-value`)
  
  smooth_table
})

# Apply FDR correction across these "by" smooths
by_smooth_pvals_nic <- by_smooth_pvals_nic %>%
  mutate(p_value_fdr = p.adjust(p_value, method = "bonferroni"))

by_smooth_pvals_nic

plot(ggpredict(fits_nic$caudate_cluster1, terms = c("age", "caudate_cluster1")))
plot(ggpredict(fits_nic$putamen_cluster1, terms = c("age", "putamen_cluster1")))
```

Yearly nicotine. Kinda weird in how it is phrased tbh.

```{r Nicotine year TVEM}
by_vars <- c(
  "putamen_cluster1",
  "putamen_cluster2",
  "caudate_cluster1",
  "caudate_cluster2",
  "nacc_cluster1",
  "nacc_cluster2"
)

# Storage lists
fits_nicy <- list()
model_warnings_nicy <- list()

for (v in by_vars) {

  # build formula
  formula <- as.formula(
    paste0(
      "cddr5 ~ 1 + sex_fac + 
       s(age, k = 5) + 
       s(age, by = ", v, ", k = 5)"
    )
  )

  # fit model
  fit <- gamm(
    formula,
    random = list(subject = ~1, subject = ~0 + age),
    data = dat_nc_cddr5,
    method = "REML",
    na.action = na.exclude
  )

  # store model
  fits_nicy[[v]] <- fit

  # capture warnings produced during summary()
  w <- NULL
  invisible(
    withCallingHandlers(
      summary(fit$gam),
      warning = function(wrn) {
        w <<- c(w, conditionMessage(wrn))
        invokeRestart("muffleWarning")
      }
    )
  )

  model_warnings_nicy[[v]] <- w
}

# Optional: print GAM summaries
lapply(fits_nicy, function(x) summary(x$gam))
model_warnings_nicy

#Explicity showing pvalues

# Extract only "by" smooth p-values across all models
by_smooth_pvals_nicy <- map_df(seq_along(fits_nicy), function(i) {
  s <- summary(fits_nicy[[i]]$gam)
  
  # Keep only smooths where the term name contains "by="
  smooth_table <- as.data.frame(s$s.table) %>%
    mutate(term = rownames(s$s.table),
           model = i) %>%
    filter(grepl(":", term)) %>%
    select(model, term, edf = edf, F, p_value = `p-value`)
  
  smooth_table
})

# Apply FDR correction across these "by" smooths
by_smooth_pvals_nicy <- by_smooth_pvals_nicy %>%
  mutate(p_value_fdr = p.adjust(p_value, method = "bonferroni"))

by_smooth_pvals_nicy

plot(ggpredict(fits_nicy$caudate_cluster1, terms = c("age", "caudate_cluster1")))
plot(ggpredict(fits_nicy$putamen_cluster1, terms = c("age", "putamen_cluster1")))
```

Checking yearly cannabis.

```{r Cannabis year TVEM}
by_vars <- c(
  "putamen_cluster1",
  "putamen_cluster2",
  "caudate_cluster1",
  "caudate_cluster2",
  "nacc_cluster1",
  "nacc_cluster2"
)

# Storage lists
fits_can <- list()
model_warnings_can <- list()

for (v in by_vars) {

  # build formula
  formula <- as.formula(
    paste0(
      "cddr56 ~ 1 + sex_fac + 
       s(age, k = 5) + 
       s(age, by = ", v, ", k = 5)"
    )
  )

  # fit model
  fit <- gamm(
    formula,
    random = list(subject = ~1, subject = ~0 + age),
    data = dat_nc,
    method = "REML",
    na.action = na.exclude
  )

  # store model
  fits_can[[v]] <- fit

  # capture warnings produced during summary()
  w <- NULL
  invisible(
    withCallingHandlers(
      summary(fit$gam),
      warning = function(wrn) {
        w <<- c(w, conditionMessage(wrn))
        invokeRestart("muffleWarning")
      }
    )
  )

  model_warnings_can[[v]] <- w
}

# Optional: print GAM summaries
lapply(fits_can, function(x) summary(x$gam))
model_warnings_can

#Explicity showing pvalues

# Extract only "by" smooth p-values across all models
by_smooth_pvals_can <- map_df(seq_along(fits_can), function(i) {
  s <- summary(fits_can[[i]]$gam)
  
  # Keep only smooths where the term name contains "by="
  smooth_table <- as.data.frame(s$s.table) %>%
    mutate(term = rownames(s$s.table),
           model = i) %>%
    filter(grepl(":", term)) %>%
    select(model, term, edf = edf, F, p_value = `p-value`)
  
  smooth_table
})

# Apply FDR correction across these "by" smooths
by_smooth_pvals_can <- by_smooth_pvals_can %>%
  mutate(p_value_fdr = p.adjust(p_value, method = "bonferroni"))

by_smooth_pvals_can


plot(ggpredict(fits_can$caudate_cluster1, terms = c("age", "caudate_cluster1")))
plot(ggpredict(fits_can$putamen_cluster1, terms = c("age", "putamen_cluster1")))
```


Checkinig monthly cannabis

```{r Cannabis month TVEM}
by_vars <- c(
  "putamen_cluster1",
  "putamen_cluster2",
  "caudate_cluster1",
  "caudate_cluster2",
  "nacc_cluster1",
  "nacc_cluster2"
)

# Storage lists
fits_canm <- list()
model_warnings_canm <- list()

for (v in by_vars) {

  # build formula
  formula <- as.formula(
    paste0(
      "cddr57 ~ 1 + sex_fac + 
       s(age, k = 5) + 
       s(age, by = ", v, ", k = 5)"
    )
  )

  # fit model
  fit <- gamm(
    formula,
    random = list(subject = ~1, subject = ~0 + age),
    data = dat_nc,
    method = "REML",
    na.action = na.exclude
  )

  # store model
  fits_canm[[v]] <- fit

  # capture warnings produced during summary()
  w <- NULL
  invisible(
    withCallingHandlers(
      summary(fit$gam),
      warning = function(wrn) {
        w <<- c(w, conditionMessage(wrn))
        invokeRestart("muffleWarning")
      }
    )
  )

  model_warnings_canm[[v]] <- w
}

# Optional: print GAM summaries
lapply(fits_canm, function(x) summary(x$gam))
model_warnings_canm

#Explicity showing pvalues

# Extract only "by" smooth p-values across all models
by_smooth_pvals_canm <- map_df(seq_along(fits_canm), function(i) {
  s <- summary(fits_canm[[i]]$gam)
  
  # Keep only smooths where the term name contains "by="
  smooth_table <- as.data.frame(s$s.table) %>%
    mutate(term = rownames(s$s.table),
           model = i) %>%
    filter(grepl(":", term)) %>%
    select(model, term, edf = edf, F, p_value = `p-value`)
  
  smooth_table
})

# Apply FDR correction across these "by" smooths
by_smooth_pvals_canm <- by_smooth_pvals_canm %>%
  mutate(p_value_fdr = p.adjust(p_value, method = "bonferroni"))

by_smooth_pvals_canm


plot(ggpredict(fits_canm$caudate_cluster1, terms = c("age", "caudate_cluster1")))
plot(ggpredict(fits_canm$putamen_cluster1, terms = c("age", "putamen_cluster1")))
```


TVEM analyses do not replicate across substances.

# Initiataion analyses

```{r Initiation cleaning}
#Merging with long data
dat_nc_long <- merge(dat_long, nc_dat, by = c("subject", "visit"))

#Excplicity testing via anova-like approach
dat_nc_long <- dat_nc_long %>%
  mutate(seed_cluster_fac = factor(seed_cluster,
                                   levels = c("nacc_cluster1", "nacc_cluster2", 
                                              "caudate_cluster1", "caudate_cluster2", 
                                              "putamen_cluster1", "putamen_cluster2")))

#Smoking onset? Before/after/never smoke
dat_nc_long_ba <- dat_nc_long %>%
  group_by(subject) %>%
  mutate(
    nicotine_onset = ifelse(any(cddr3 > 0), min(age[cddr3 > 0], na.rm = TRUE), NA),
    nicotine_phase = case_when(
      is.na(nicotine_onset) ~ "Never",   # Never drank
      age < nicotine_onset  ~ "Before",  # Before first drinking report
      age >= nicotine_onset ~ "After"    # After first drinking report
    )
  ) %>%
  ungroup() %>%
  mutate(nicotine_phase = factor(nicotine_phase, levels = c("Before", "After", "Never")),
         seed_fac = factor(seed_fac),
         subject_fac = factor(subject_fac))

#Cannabis onset. Before/after/never smoke.
dat_nc_long_ba <- dat_nc_long_ba %>%
  group_by(subject) %>%
  mutate(
    cannabis_onset = ifelse(any(cddr56 > 0), min(age[cddr56 > 0], na.rm = TRUE), NA),
    cannabis_phase = case_when(
      is.na(cannabis_onset) ~ "Never",   # Never drank
      age < cannabis_onset  ~ "Before",  # Before first drinking report
      age >= cannabis_onset ~ "After"    # After first drinking report
    )
  ) %>%
  ungroup() %>%
  mutate(cannabis_phase = factor(cannabis_phase, levels = c("Before", "After", "Never")),
         seed_fac = factor(seed_fac),
         subject_fac = factor(subject_fac))

#Cannabis onset month. Before/after/never smoke.
dat_nc_long_ba <- dat_nc_long_ba %>%
  group_by(subject) %>%
  mutate(
    cannabis_onsetm = ifelse(any(cddr57 > 0), min(age[cddr57 > 0], na.rm = TRUE), NA),
    cannabis_phasem = case_when(
      is.na(cannabis_onsetm) ~ "Never",   # Never drank
      age < cannabis_onsetm  ~ "Before",  # Before first drinking report
      age >= cannabis_onsetm ~ "After"    # After first drinking report
    )
  ) %>%
  ungroup() %>%
  mutate(cannabis_phasem = factor(cannabis_phasem, levels = c("Before", "After", "Never")),
         seed_fac = factor(seed_fac),
         subject_fac = factor(subject_fac))
```

Now I should test this across nicotine monthly, and cannabis monthly and yearly.

This should result in around 4 plots per substance variable, so around 12ish plots? 
From there I am probably ok with just putting it all in one giant figure? I dunno yet. 

```{r Initiation cddr 3 monthly nicotine}
nic_init <- bam(mean_value ~ 1 + sex_fac  +  seed_cluster_fac*nicotine_phase +
                                s(age, k = 3) + 
                                s(subject_fac, bs = 're') +
                                s(subject_fac, age, bs = 're'), 
                   data = dat_nc_long_ba, 
                   na.action = na.exclude, 
                   method = "fREML")
summary(nic_init)

plot(ggpredict(nic_init, terms = c("nicotine_phase", 
                                      "seed_cluster_fac")))

#Probing the intereactions
emm_results_nic <- emmeans(nic_init, ~ nicotine_phase | seed_cluster_fac)
contrasts_nic <- contrast(emm_results_nic, method = "pairwise", adjust = "bonferroni")
contrasts_nic

#Just doing adjustment manually.
contrasts_dat_nic <- as.data.frame(contrasts_nic)
contrasts_dat_nic$p_adj <- p.adjust(contrasts_dat_nic$p.value, method = "bonferroni")
#making sig column for easy viewing.
contrasts_dat_nic$sig <- contrasts_dat_nic$p_adj < 0.05

#I like above approach, where I just manually correct for the 16 tests.

plot(emm_results_nic, horizontal = TRUE)
plot(contrasts_nic)

q_nic <- as.data.frame(emm_results_nic)
```

Seems to be a similar effect as alcohol freq.

```{r PLOTS Initiation cddr 3 monthly nicotine}
seed_phase_nic <- ggplot(q_nic %>% 
                       filter(nicotine_phase %in% c("Before", "After", "Never")), 
                     aes(x = seed_cluster_fac, 
                         y = emmean, 
                         group = nicotine_phase,
                         color = seed_cluster_fac)) + 
  geom_point(position = position_dodge(width = 0.7), size = 1) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.2, 
                position = position_dodge(width = 0.7)) +
  scale_color_manual(values = c("putamen_cluster1" = "#1b9e77",
                                "putamen_cluster2" = "#1b3877ff",
                                "caudate_cluster1" = "#7570b3",
                                "caudate_cluster2" = "#7500b3ff",
                                "nacc_cluster1" = "#d95f02",
                                "nacc_cluster2" = "#e6ab02"
                                )) +
  scale_x_discrete(labels = rep(c("Before    After   Never"), 6)) +
  annotate("text", x = 1:6, y = .20, 
           label = c("NAcc Cluster 1", "NAcc Cluster 2", 
                     "Caudate Cluster 1", "Caudate Cluster 2",
                     "Putamen Cluster 1", "Putamen Cluster 2"),
           vjust = -0.5, hjust = 0.5, size = 2) +
  labs(x = "Nicotine Phase", y = "Estimated Mean") +
  theme_modern() +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    legend.position = "none"
    )
seed_phase_nic

ggsave(plot = seed_phase_nic,
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/nicotine/seed_phase.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 160, 
       height = 60)

#I also want to isolate figures that I will carry over. Putamen 1
seed_phase_nicotine_p1 <- ggplot(q_nic %>% 
                       filter(nicotine_phase %in% c("Before", "After", "Never")) %>%
                       filter(seed_cluster_fac == "putamen_cluster1"), 
                     aes(x = seed_cluster_fac, 
                         y = emmean, 
                         group = nicotine_phase,
                         color = seed_cluster_fac)) + 
  geom_point(position = position_dodge(width = 0.7), size = 1) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.2, 
                position = position_dodge(width = 0.7)) +
  scale_color_manual(values = c("putamen_cluster1" = "#1b9e77")) +
  scale_y_continuous(limits = c(-0.2, 0.2),
                     breaks = seq(-0.2,0.2, by = 0.1)) +
  #scale_x_discrete(labels = rep(c("Before    After   Never"), 1)) +
  #annotate("text", x = 0, y = .20, 
  #         label = c("Putamen Cluster 1"),
  #         vjust = -0.5, hjust = 0.5, size = 2) +
  labs(x = "", y = "") +
  theme_modern() +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    legend.position = "none"
    )
seed_phase_nicotine_p1

ggsave(plot = seed_phase_nicotine_p1,
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/nicotine/putamen1_init.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 36, 
       height = 60)

#I also want to isolate figures that I will carry over. caudate 2
seed_phase_nicotine_c2 <- ggplot(q_nic %>% 
                       filter(nicotine_phase %in% c("Before", "After", "Never")) %>%
                       filter(seed_cluster_fac == "caudate_cluster2"), 
                     aes(x = seed_cluster_fac, 
                         y = emmean, 
                         group = nicotine_phase,
                         color = seed_cluster_fac)) + 
  geom_point(position = position_dodge(width = 0.7), size = 1) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.2, 
                position = position_dodge(width = 0.7)) +
  scale_color_manual(values = c("caudate_cluster2" = "#7500b3ff")) +
  scale_y_continuous(limits = c(-0.2, 0.3),
                     breaks = seq(-0.2,0.3, by = 0.1)) +
  #scale_x_discrete(labels = rep(c("Before    After   Never"), 1)) +
  #annotate("text", x = 0, y = .20, 
  #         label = c("Putamen Cluster 1"),
  #         vjust = -0.5, hjust = 0.5, size = 2) +
  labs(x = "", y = "") +
  theme_modern() +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    legend.position = "none"
    )
seed_phase_nicotine_c2

ggsave(plot = seed_phase_nicotine_c2,
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/nicotine/caudate2_init.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 36, 
       height = 60)

#I also want to isolate figures that I will carry over. nacc 1
seed_phase_nicotine_n1 <- ggplot(q_nic %>% 
                       filter(nicotine_phase %in% c("Before", "After", "Never")) %>%
                       filter(seed_cluster_fac == "nacc_cluster1"), 
                     aes(x = seed_cluster_fac, 
                         y = emmean, 
                         group = nicotine_phase,
                         color = seed_cluster_fac)) + 
  geom_point(position = position_dodge(width = 0.7), size = 1) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.2, 
                position = position_dodge(width = 0.7)) +
  scale_color_manual(values = c("nacc_cluster1" = "#d95f02")) +
  scale_y_continuous(limits = c(-0.2, 0.2),
                     breaks = seq(-0.2,0.2, by = 0.1)) +
  #scale_x_discrete(labels = rep(c("Before    After   Never"), 1)) +
  #annotate("text", x = 0, y = .20, 
  #         label = c("Putamen Cluster 1"),
  #         vjust = -0.5, hjust = 0.5, size = 2) +
  labs(x = "", y = "") +
  theme_modern() +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    legend.position = "none"
    )
seed_phase_nicotine_n1

ggsave(plot = seed_phase_nicotine_n1,
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/nicotine/nacc1_init.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 36, 
       height = 60)
```


Cannabis

```{r Initiation cddr 56 yearly cannabis}
cana_init <- bam(mean_value ~ 1 + sex_fac  +  seed_cluster_fac*cannabis_phase +
                                s(age, k = 3) + 
                                s(subject_fac, bs = 're') +
                                s(subject_fac, age, bs = 're'), 
                   data = dat_nc_long_ba, 
                   na.action = na.exclude, 
                   method = "fREML")
summary(cana_init)

plot(ggpredict(cana_init, terms = c("seed_cluster_fac", 
                                      "cannabis_phase")))

plot(ggpredict(cana_init, terms = c("age", "seed_cluster_fac")))

#Probing the intereactions
emm_results_cana <- emmeans(cana_init, ~ cannabis_phase | seed_cluster_fac)
contrasts_cana <- contrast(emm_results_cana, method = "pairwise", adjust = "bonferroni")
contrasts_cana

#Just doing adjustment manually.
contrasts_dat_cana <- as.data.frame(contrasts_cana)
contrasts_dat_cana$p_adj <- p.adjust(contrasts_dat_cana$p.value, method = "bonferroni")
#making sig column for easy viewing.
contrasts_dat_cana$sig <- contrasts_dat_cana$p_adj < 0.05

#I like above approach, where I just manually correct for the 16 tests.

plot(emm_results_cana, horizontal = TRUE)
plot(contrasts_cana)

q_cana <- as.data.frame(emm_results_cana)
```

Plots for yearly cannabis use.

```{r PLOTS Initiation cddr 56 yearly cannabis}
seed_phase_cannabis <- ggplot(q_cana %>% 
                       filter(cannabis_phase %in% c("Before", "After", "Never")), 
                     aes(x = seed_cluster_fac, 
                         y = emmean, 
                         group = cannabis_phase,
                         color = seed_cluster_fac)) + 
  geom_point(position = position_dodge(width = 0.7), size = 1) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.2, 
                position = position_dodge(width = 0.7)) +
  scale_color_manual(values = c("putamen_cluster1" = "#1b9e77",
                                "putamen_cluster2" = "#1b3877ff",
                                "caudate_cluster1" = "#7570b3",
                                "caudate_cluster2" = "#7500b3ff",
                                "nacc_cluster1" = "#d95f02",
                                "nacc_cluster2" = "#e6ab02"
                                )) +
  scale_x_discrete(labels = rep(c("Before    After   Never"), 6)) +
  annotate("text", x = 1:6, y = .15, 
           label = c("NAcc Cluster 1", "NAcc Cluster 2", 
                     "Caudate Cluster 1", "Caudate Cluster 2",
                     "Putamen Cluster 1", "Putamen Cluster 2"),
           vjust = -0.5, hjust = 0.5, size = 2) +
  labs(x = "Cannabis Phase", y = "Estimated Mean") +
  theme_modern() +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    legend.position = "none"
    )
seed_phase_cannabis

ggsave(plot = seed_phase_cannabis,
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/cannabis/yearly/seed_phase_cannabis.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 160, 
       height = 60)

#I also want to isolate figures that I will carry over. Putamen 1
seed_phase_cannabis_p1 <- ggplot(q_cana %>% 
                       filter(cannabis_phase %in% c("Before", "After", "Never")) %>%
                       filter(seed_cluster_fac == "putamen_cluster1"), 
                     aes(x = seed_cluster_fac, 
                         y = emmean, 
                         group = cannabis_phase,
                         color = seed_cluster_fac)) + 
  geom_point(position = position_dodge(width = 0.7), size = 1) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.2, 
                position = position_dodge(width = 0.7)) +
  scale_color_manual(values = c("putamen_cluster1" = "#1b9e77")) +
  scale_y_continuous(limits = c(-0.2, 0.2),
                     breaks = seq(-0.2,0.2, by = 0.1)) +
  #scale_x_discrete(labels = rep(c("Before    After   Never"), 1)) +
  #annotate("text", x = 0, y = .20, 
  #         label = c("Putamen Cluster 1"),
  #         vjust = -0.5, hjust = 0.5, size = 2) +
  labs(x = "", y = "") +
  theme_modern() +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    legend.position = "none"
    )
seed_phase_cannabis_p1

ggsave(plot = seed_phase_cannabis_p1,
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/cannabis/yearly/putamen1_init.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 36, 
       height = 60)

#I also want to isolate figures that I will carry over. caudate 2
seed_phase_cannabis_c2 <- ggplot(q_cana %>% 
                       filter(cannabis_phase %in% c("Before", "After", "Never")) %>%
                       filter(seed_cluster_fac == "caudate_cluster2"), 
                     aes(x = seed_cluster_fac, 
                         y = emmean, 
                         group = cannabis_phase,
                         color = seed_cluster_fac)) + 
  geom_point(position = position_dodge(width = 0.7), size = 1) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.2, 
                position = position_dodge(width = 0.7)) +
  scale_color_manual(values = c("caudate_cluster2" = "#7500b3ff")) +
  scale_y_continuous(limits = c(-0.2, 0.2),
                     breaks = seq(-0.2,0.2, by = 0.1)) +
  #scale_x_discrete(labels = rep(c("Before    After   Never"), 1)) +
  #annotate("text", x = 0, y = .20, 
  #         label = c("Putamen Cluster 1"),
  #         vjust = -0.5, hjust = 0.5, size = 2) +
  labs(x = "", y = "") +
  theme_modern() +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    legend.position = "none"
    )
seed_phase_cannabis_c2

ggsave(plot = seed_phase_cannabis_c2,
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/cannabis/yearly/caudate2_init.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 36, 
       height = 60)

#I also want to isolate figures that I will carry over. nacc 1
seed_phase_cannabis_n1 <- ggplot(q_cana %>% 
                       filter(cannabis_phase %in% c("Before", "After", "Never")) %>%
                       filter(seed_cluster_fac == "nacc_cluster1"), 
                     aes(x = seed_cluster_fac, 
                         y = emmean, 
                         group = cannabis_phase,
                         color = seed_cluster_fac)) + 
  geom_point(position = position_dodge(width = 0.7), size = 1) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.2, 
                position = position_dodge(width = 0.7)) +
  scale_color_manual(values = c("nacc_cluster1" = "#d95f02")) +
  scale_y_continuous(limits = c(-0.2, 0.2),
                     breaks = seq(-0.2,0.2, by = 0.1)) +
  #scale_x_discrete(labels = rep(c("Before    After   Never"), 1)) +
  #annotate("text", x = 0, y = .20, 
  #         label = c("Putamen Cluster 1"),
  #         vjust = -0.5, hjust = 0.5, size = 2) +
  labs(x = "", y = "") +
  theme_modern() +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    legend.position = "none"
    )
seed_phase_cannabis_n1

ggsave(plot = seed_phase_cannabis_n1,
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/cannabis/yearly/nacc1_init.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 36, 
       height = 60)
```

Cannabis month

```{r Initiation cddr 57 monthly cannabis}
cana_initm <- bam(mean_value ~ 1 + sex_fac  +  seed_cluster_fac*cannabis_phasem +
                                s(age, k = 3) + 
                                s(subject_fac, bs = 're') +
                                s(subject_fac, age, bs = 're'), 
                   data = dat_nc_long_ba, 
                   na.action = na.exclude, 
                   method = "fREML")
summary(cana_initm)

plot(ggpredict(cana_initm, terms = c("seed_cluster_fac", 
                                      "cannabis_phasem")))

plot(ggpredict(cana_initm, terms = c("age", "seed_cluster_fac")))

#Probing the intereactions
emm_results_canam <- emmeans(cana_initm, ~ cannabis_phasem | seed_cluster_fac)
contrasts_canam <- contrast(emm_results_canam, method = "pairwise", adjust = "bonferroni")
contrasts_canam

#Just doing adjustment manually.
contrasts_dat_canam <- as.data.frame(contrasts_canam)
contrasts_dat_canam$p_adj <- p.adjust(contrasts_dat_canam$p.value, method = "bonferroni")
#making sig column for easy viewing.
contrasts_dat_canam$sig <- contrasts_dat_canam$p_adj < 0.05

#I like above approach, where I just manually correct for the 16 tests.

plot(emm_results_canam, horizontal = TRUE)
plot(contrasts_canam)

q_canam <- as.data.frame(emm_results_canam)
```


```{r PLOTS Initiation cddr 57 monthly cannabis}
seed_phase_cannabism <- ggplot(q_canam %>% 
                       filter(cannabis_phasem %in% c("Before", "After", "Never")), 
                     aes(x = seed_cluster_fac, 
                         y = emmean, 
                         group = cannabis_phasem,
                         color = seed_cluster_fac)) + 
  geom_point(position = position_dodge(width = 0.7), size = 1) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.2, 
                position = position_dodge(width = 0.7)) +
  scale_color_manual(values = c("putamen_cluster1" = "#1b9e77",
                                "putamen_cluster2" = "#1b3877ff",
                                "caudate_cluster1" = "#7570b3",
                                "caudate_cluster2" = "#7500b3ff",
                                "nacc_cluster1" = "#d95f02",
                                "nacc_cluster2" = "#e6ab02"
                                )) +
  scale_x_discrete(labels = rep(c("Before    After   Never"), 6)) +
  annotate("text", x = 1:6, y = .20, 
           label = c("NAcc Cluster 1", "NAcc Cluster 2", 
                     "Caudate Cluster 1", "Caudate Cluster 2",
                     "Putamen Cluster 1", "Putamen Cluster 2"),
           vjust = -0.5, hjust = 0.5, size = 2) +
  labs(x = "Cannabis Phase", y = "Estimated Mean") +
  theme_modern() +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    legend.position = "none"
    )
seed_phase_cannabism

ggsave(plot = seed_phase_cannabism,
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/cannabis/monthly/seed_phase.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 160, 
       height = 60)

#I also want to isolate figures that I will carry over. Putamen 1
seed_phase_cannabism_p1 <- ggplot(q_canam %>% 
                       filter(cannabis_phasem %in% c("Before", "After", "Never")) %>%
                       filter(seed_cluster_fac == "putamen_cluster1"), 
                     aes(x = seed_cluster_fac, 
                         y = emmean, 
                         group = cannabis_phasem,
                         color = seed_cluster_fac)) + 
  geom_point(position = position_dodge(width = 0.7), size = 1) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.2, 
                position = position_dodge(width = 0.7)) +
  scale_color_manual(values = c("putamen_cluster1" = "#1b9e77")) +
  scale_y_continuous(limits = c(-0.2, 0.2),
                     breaks = seq(-0.2,0.2, by = 0.1)) +
  #scale_x_discrete(labels = rep(c("Before    After   Never"), 1)) +
  #annotate("text", x = 0, y = .20, 
  #         label = c("Putamen Cluster 1"),
  #         vjust = -0.5, hjust = 0.5, size = 2) +
  labs(x = "", y = "") +
  theme_modern() +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    legend.position = "none"
    )
seed_phase_cannabism_p1

ggsave(plot = seed_phase_cannabism_p1,
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/cannabis/monthly/putamen1_init.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 36, 
       height = 60)

#I also want to isolate figures that I will carry over. caudate 2
seed_phase_cannabism_c2 <- ggplot(q_canam %>% 
                       filter(cannabis_phasem %in% c("Before", "After", "Never")) %>%
                       filter(seed_cluster_fac == "caudate_cluster2"), 
                     aes(x = seed_cluster_fac, 
                         y = emmean, 
                         group = cannabis_phasem,
                         color = seed_cluster_fac)) + 
  geom_point(position = position_dodge(width = 0.7), size = 1) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.2, 
                position = position_dodge(width = 0.7)) +
  scale_color_manual(values = c("caudate_cluster2" = "#7500b3ff")) +
  scale_y_continuous(limits = c(-0.2, 0.2),
                     breaks = seq(-0.2,0.2, by = 0.1)) +
  #scale_x_discrete(labels = rep(c("Before    After   Never"), 1)) +
  #annotate("text", x = 0, y = .20, 
  #         label = c("Putamen Cluster 1"),
  #         vjust = -0.5, hjust = 0.5, size = 2) +
  labs(x = "", y = "") +
  theme_modern() +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    legend.position = "none"
    )
seed_phase_cannabism_c2

ggsave(plot = seed_phase_cannabism_c2,
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/cannabis/monthly/caudate2_init.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 36, 
       height = 60)

#I also want to isolate figures that I will carry over. nacc 1
seed_phase_cannabism_n1 <- ggplot(q_canam %>% 
                       filter(cannabis_phasem %in% c("Before", "After", "Never")) %>%
                       filter(seed_cluster_fac == "nacc_cluster1"), 
                     aes(x = seed_cluster_fac, 
                         y = emmean, 
                         group = cannabis_phasem,
                         color = seed_cluster_fac)) + 
  geom_point(position = position_dodge(width = 0.7), size = 1) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), 
                width = 0.2, 
                position = position_dodge(width = 0.7)) +
  scale_color_manual(values = c("nacc_cluster1" = "#d95f02")) +
  scale_y_continuous(limits = c(-0.2, 0.2),
                     breaks = seq(-0.2,0.2, by = 0.1)) +
  #scale_x_discrete(labels = rep(c("Before    After   Never"), 1)) +
  #annotate("text", x = 0, y = .20, 
  #         label = c("Putamen Cluster 1"),
  #         vjust = -0.5, hjust = 0.5, size = 2) +
  labs(x = "", y = "") +
  theme_modern() +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    #legend.text = element_text(size = 5),
    #legend.title = element_text(size = 6),
    #legend.key.size = unit(0.3, "cm"),
    legend.position = "none"
    )
seed_phase_cannabism_n1

ggsave(plot = seed_phase_cannabism_n1,
       filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/cannabis/monthly/nacc1_init.pdf", 
       device = "pdf", 
       dpi = 300,
       units = "mm",
       width = 36, 
       height = 60)
```


# Analyses and Figure showing growth model colored by drinking onset (nacc)

Monthly Nicotine

```{r Growth Initiation monthly nicotine analysis and plot}
#dat_long_ba$time_since_onset <- dat_long_ba$age - dat_long_ba$drinking_onset

#Creating new dataframe
dat_pred_nc_nacc1 <- dat_nc_long_ba %>%
  filter(seed_cluster == "nacc_cluster1") 

#Seeing if additional grouping variable works
dat_pred_nc_nacc1 <- dat_pred_nc_nacc1 %>%
  mutate(nicotine_group = ifelse(nicotine_phase == "Never", "Never", "Initiated"),
         nicotine_group = factor(nicotine_group, levels = c("Initiated", "Never")))

#Oredered?
dat_pred_nc_nacc1$nicotine_group_ord <- factor(dat_pred_nc_nacc1$nicotine_group, 
                             levels = c("Never", "Initiated"), 
                             ordered = TRUE)

test <- bam(mean_value ~ 1 + sex_fac + nicotine_group_ord +
              s(age, k = 5) +
              s(age, by = nicotine_group_ord, k = 5) +
              s(subject_fac, bs = 're') +
              s(age, subject_fac, bs = 're'), 
            data = dat_pred_nc_nacc1, 
            na.action = na.exclude, 
            method = "fREML")

summary(test)
#Function to calculate difference smooths
#test_diff <- difference_smooths(test, select = c("s(age)", "s(age):drinking_group_ordInitiated"), partial_match = FALSE)

#draw(test_diff)

#Gratia plot
draw(test)
#mgcv plot
plot(test, pages = 1)

plot(ggpredict(test, terms = c("age")))
plot(ggpredict(test, terms = c("nicotine_group_ord")))
plot(ggpredict(test, terms = c("age", "nicotine_group_ord")), show_ci = FALSE)


# Add predicted values to your data

dat_pred_nc_nacc1$predicted <-  predict(test, newdata = dat_pred_nc_nacc1, type = "response")




#If I want to color lines accordingly, I need to create average age by drinking-phase.
cutoff_age <- dat_pred_nc_nacc1 %>%
  filter(nicotine_phase == "After") %>%
  group_by(subject) %>%
  summarize(min_after_age = min(age), .groups = "drop") %>%
  summarize(avg_cutoff = mean(min_after_age, na.rm = TRUE)) %>%
  pull(avg_cutoff)

dat_pred_nc_nacc1 <- dat_pred_nc_nacc1 %>%
  mutate(phase_age = case_when(
    nicotine_phase == "Never" ~ "Never",
    age < cutoff_age ~ "Before",
    age >= cutoff_age ~ "After"
  ))

nacc1 <- ggplot(dat_pred_nc_nacc1, aes(x = age, y = predicted, group = subject, color = nicotine_phase)) +
  #geom_point(alpha = 0.2, size = 0.05) +
  geom_line(alpha = 0.2, size = 0.1) +
      scale_color_manual(values = c(
    "After" = "#d95f02ff",
    "Before" = "black", 
    "Never" = "black"))   +   
    new_scale_color() +
  new_scale_fill() +
  geom_smooth(aes(group = phase_age, color = phase_age), size = 0.3, show.legend = FALSE) +
  labs(y = "", x = "") +
    scale_color_manual(values = c(
    "After" = "#d95f02ff",  
    "Before" = "black", 
    "Never" = "black"
  )) +
    facet_wrap(~nicotine_group, nrow = 1) + 
 # geom_vline(xintercept = cutoff_age, linetype = "dashed", size = 0.2) +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    legend.position = "none",
    strip.text = element_blank(),
    panel.background = element_blank(),   # remove gray panel
    panel.grid = element_blank(),         # remove grid lines
    plot.background = element_blank()     # remove outer gray
  )

nacc1

ggsave(filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/nicotine/NAcc_pred_ban_plot.pdf", 
      plot = nacc1,
       device = "pdf", 
       dpi = 300, 
      units = "mm",
       width = 90, 
       height = 60)

#Caudate


# Add predicted values to your data
dat_nc_long_ba_caud2 <- dat_nc_long_ba %>%
  filter(seed_cluster == "caudate_cluster2") 

#Seeing if additional grouping variable works
dat_nc_long_ba_caud2<- dat_nc_long_ba_caud2 %>%
  mutate(nicotine_group = ifelse(nicotine_phase == "Never", "Never", "Initiated"),
         nicotine_group = factor(nicotine_group, levels = c("Initiated", "Never")))

#Oredered?
dat_nc_long_ba_caud2$nicotine_group_ord <- factor(dat_nc_long_ba_caud2$nicotine_group, 
                             levels = c("Never", "Initiated"), 
                             ordered = TRUE)

#Seeing if there are actual trajectory differences
test2 <- bam(mean_value ~ 1 + sex_fac  +  nicotine_group_ord +
              s(age, k = 5) +
              s(age, by = nicotine_group_ord, k = 5) +
              s(subject_fac, bs = 're') +
              s(age, subject_fac, bs = 're'), 
            data = dat_nc_long_ba_caud2, 
            na.action = na.exclude, 
            method = "fREML")

summary(test2)
draw(test2)
plot(ggpredict(test2, terms = c("age")))
plot(ggpredict(test2, terms = c("nicotine_group_ord")))
plot(ggpredict(test2, terms = c("age", "nicotine_group_ord")))

#Preds
dat_nc_long_ba_caud2$predicted <-  predict(test2, newdata = dat_nc_long_ba_caud2, type = "response")




#If I want to color lines accordingly, I need to create average age by drinking-phase.
cutoff_age <- dat_nc_long_ba_caud2 %>%
  filter(nicotine_phase == "After") %>%
  group_by(subject) %>%
  summarize(min_after_age = min(age), .groups = "drop") %>%
  summarize(avg_cutoff = mean(min_after_age, na.rm = TRUE)) %>%
  pull(avg_cutoff)

dat_nc_long_ba_caud2 <- dat_nc_long_ba_caud2 %>%
  mutate(phase_age = case_when(
   nicotine_phase == "Never" ~ "Never",
    age < cutoff_age ~ "Before",
    age >= cutoff_age ~ "After"
  ))

caud2 <- ggplot(dat_nc_long_ba_caud2, aes(x = age, y = predicted, group = subject, color = nicotine_phase)) +
  #geom_point(alpha = 0.2, size = 0.05) +
  geom_line(alpha = 0.2, size = 0.1) +
      scale_color_manual(values = c(
    "After" = "#7500b3ff",
    "Before" = "black", 
    "Never" = "black"))   +   
    new_scale_color() +
  new_scale_fill() +
  geom_smooth(aes(group = phase_age, color = phase_age), size = 0.3, show.legend = FALSE) +
  labs(y = "", x = "") +
    scale_color_manual(values = c(
    "After" = "#7500b3ff",  
    "Before" = "black", 
    "Never" = "black"
  )) +
    facet_wrap(~nicotine_group, nrow = 1) + 
 # geom_vline(xintercept = cutoff_age, linetype = "dashed", size = 0.2) +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    legend.position = "none",
    strip.text = element_blank(),
    panel.background = element_blank(),   # remove gray panel
    panel.grid = element_blank(),         # remove grid lines
    plot.background = element_blank()     # remove outer gray
  )

caud2

ggsave(filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/nicotine/caud_pred_ban_plot.pdf", 
      plot = caud2,
       device = "pdf", 
       dpi = 300, 
      units = "mm",
       width = 90, 
       height = 60)
```

Yearly Cannabis

```{r Growth Initiation yearly cannabis analysis and plot}
#dat_long_ba$time_since_onset <- dat_long_ba$age - dat_long_ba$drinking_onset

#Creating new dataframe
#dat_pred_nc_nacc1 <- dat_nc_long_ba %>%
#  filter(seed_cluster == "nacc_cluster1") 

#Seeing if additional grouping variable works
dat_pred_nc_nacc1 <- dat_pred_nc_nacc1 %>%
  mutate(cannabis_group = ifelse(cannabis_phase == "Never", "Never", "Initiated"),
         cannabis_group = factor(cannabis_group, levels = c("Initiated", "Never")))

#Oredered?
dat_pred_nc_nacc1$cannabis_group_ord <- factor(dat_pred_nc_nacc1$cannabis_group, 
                             levels = c("Never", "Initiated"), 
                             ordered = TRUE)

test <- bam(mean_value ~ 1 + sex_fac + cannabis_group_ord +
              s(age, k = 5) +
              s(age, by = cannabis_group_ord, k = 5) +
              s(subject_fac, bs = 're') +
              s(age, subject_fac, bs = 're'), 
            data = dat_pred_nc_nacc1, 
            na.action = na.exclude, 
            method = "fREML")

summary(test)
#Function to calculate difference smooths
#test_diff <- difference_smooths(test, select = c("s(age)", "s(age):drinking_group_ordInitiated"), partial_match = FALSE)

#draw(test_diff)

#Gratia plot
draw(test)
#mgcv plot
plot(test, pages = 1)

plot(ggpredict(test, terms = c("age")))
plot(ggpredict(test, terms = c("cannabis_group_ord")))
plot(ggpredict(test, terms = c("age", "cannabis_group_ord")), show_ci = FALSE)


# Add predicted values to your data

dat_pred_nc_nacc1$predicted <-  predict(test, newdata = dat_pred_nc_nacc1, type = "response")




#If I want to color lines accordingly, I need to create average age by drinking-phase.
cutoff_age <- dat_pred_nc_nacc1 %>%
  filter(cannabis_phase == "After") %>%
  group_by(subject) %>%
  summarize(min_after_age = min(age), .groups = "drop") %>%
  summarize(avg_cutoff = mean(min_after_age, na.rm = TRUE)) %>%
  pull(avg_cutoff)

dat_pred_nc_nacc1 <- dat_pred_nc_nacc1 %>%
  mutate(phase_age = case_when(
    cannabis_phase == "Never" ~ "Never",
    age < cutoff_age ~ "Before",
    age >= cutoff_age ~ "After"
  ))

nacc1 <- ggplot(dat_pred_nc_nacc1, aes(x = age, y = predicted, group = subject, color = cannabis_phase)) +
  #geom_point(alpha = 0.2, size = 0.05) +
  geom_line(alpha = 0.2, size = 0.1) +
      scale_color_manual(values = c(
    "After" = "#d95f02ff",
    "Before" = "black", 
    "Never" = "black"))   +   
    new_scale_color() +
  new_scale_fill() +
  geom_smooth(aes(group = phase_age, color = phase_age), size = 0.3, show.legend = FALSE) +
  labs(y = "", x = "") +
    scale_color_manual(values = c(
    "After" = "#d95f02ff",  
    "Before" = "black", 
    "Never" = "black"
  )) +
    facet_wrap(~cannabis_group, nrow = 1) + 
 # geom_vline(xintercept = cutoff_age, linetype = "dashed", size = 0.2) +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    legend.position = "none",
    strip.text = element_blank(),
    panel.background = element_blank(),   # remove gray panel
    panel.grid = element_blank(),         # remove grid lines
    plot.background = element_blank()     # remove outer gray
  )

nacc1

ggsave(filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/cannabis/yearly/NAcc_pred_ban_plot.pdf", 
      plot = nacc1,
       device = "pdf", 
       dpi = 300, 
      units = "mm",
       width = 90, 
       height = 60)

#Caudate


# Add predicted values to your data
dat_nc_long_ba_caud2 <- dat_nc_long_ba %>%
  filter(seed_cluster == "caudate_cluster2") 

#Seeing if additional grouping variable works
dat_nc_long_ba_caud2<- dat_nc_long_ba_caud2 %>%
  mutate(cannabis_group = ifelse(cannabis_phase == "Never", "Never", "Initiated"),
         cannabis_group = factor(cannabis_group, levels = c("Initiated", "Never")))

#Oredered?
dat_nc_long_ba_caud2$cannabis_group_ord <- factor(dat_nc_long_ba_caud2$cannabis_group, 
                             levels = c("Never", "Initiated"), 
                             ordered = TRUE)

#Seeing if there are actual trajectory differences
test2 <- bam(mean_value ~ 1 + sex_fac  +  cannabis_group_ord +
              s(age, k = 5) +
              s(age, by = cannabis_group_ord, k = 5) +
              s(subject_fac, bs = 're') +
              s(age, subject_fac, bs = 're'), 
            data = dat_nc_long_ba_caud2, 
            na.action = na.exclude, 
            method = "fREML")

summary(test2)
draw(test2)
plot(ggpredict(test2, terms = c("age")))
plot(ggpredict(test2, terms = c("cannabis_group_ord")))
plot(ggpredict(test2, terms = c("age", "cannabis_group_ord")))

#Preds
dat_nc_long_ba_caud2$predicted <-  predict(test2, newdata = dat_nc_long_ba_caud2, type = "response")




#If I want to color lines accordingly, I need to create average age by drinking-phase.
cutoff_age <- dat_nc_long_ba_caud2 %>%
  filter(cannabis_phase == "After") %>%
  group_by(subject) %>%
  summarize(min_after_age = min(age), .groups = "drop") %>%
  summarize(avg_cutoff = mean(min_after_age, na.rm = TRUE)) %>%
  pull(avg_cutoff)

dat_nc_long_ba_caud2 <- dat_nc_long_ba_caud2 %>%
  mutate(phase_age = case_when(
   cannabis_phase == "Never" ~ "Never",
    age < cutoff_age ~ "Before",
    age >= cutoff_age ~ "After"
  ))

caud2 <- ggplot(dat_nc_long_ba_caud2, aes(x = age, y = predicted, group = subject, color = cannabis_phase)) +
  #geom_point(alpha = 0.2, size = 0.05) +
  geom_line(alpha = 0.2, size = 0.1) +
      scale_color_manual(values = c(
    "After" = "#7500b3ff",
    "Before" = "black", 
    "Never" = "black"))   +   
    new_scale_color() +
  new_scale_fill() +
  geom_smooth(aes(group = phase_age, color = phase_age), size = 0.3, show.legend = FALSE) +
  labs(y = "", x = "") +
    scale_color_manual(values = c(
    "After" = "#7500b3ff",  
    "Before" = "black", 
    "Never" = "black"
  )) +
    facet_wrap(~cannabis_group, nrow = 1) + 
 # geom_vline(xintercept = cutoff_age, linetype = "dashed", size = 0.2) +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    legend.position = "none",
    strip.text = element_blank(),
    panel.background = element_blank(),   # remove gray panel
    panel.grid = element_blank(),         # remove grid lines
    plot.background = element_blank()     # remove outer gray
  )

caud2

ggsave(filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/cannabis/yearly/caud_pred_ban_plot.pdf", 
      plot = caud2,
       device = "pdf", 
       dpi = 300, 
      units = "mm",
       width = 90, 
       height = 60)
```


CAnnabis monthly

```{r Growth Initiation monthly cannabis analysis and plot}
#dat_long_ba$time_since_onset <- dat_long_ba$age - dat_long_ba$drinking_onset

#Creating new dataframe
#dat_pred_nc_nacc1 <- dat_nc_long_ba %>%
#  filter(seed_cluster == "nacc_cluster1") 

#Seeing if additional grouping variable works
dat_pred_nc_nacc1 <- dat_pred_nc_nacc1 %>%
  mutate(cannabism_group = ifelse(cannabis_phasem == "Never", "Never", "Initiated"),
         cannabism_group = factor(cannabism_group, levels = c("Initiated", "Never")))

#Oredered?
dat_pred_nc_nacc1$cannabism_group_ord <- factor(dat_pred_nc_nacc1$cannabism_group, 
                             levels = c("Never", "Initiated"), 
                             ordered = TRUE)

test <- bam(mean_value ~ 1 + sex_fac + cannabism_group_ord +
              s(age, k = 5) +
              s(age, by = cannabism_group_ord, k = 5) +
              s(subject_fac, bs = 're') +
              s(age, subject_fac, bs = 're'), 
            data = dat_pred_nc_nacc1, 
            na.action = na.exclude, 
            method = "fREML")

summary(test)
#Function to calculate difference smooths
#test_diff <- difference_smooths(test, select = c("s(age)", "s(age):drinking_group_ordInitiated"), partial_match = FALSE)

#draw(test_diff)

#Gratia plot
draw(test)
#mgcv plot
plot(test, pages = 1)

plot(ggpredict(test, terms = c("age")))
plot(ggpredict(test, terms = c("cannabism_group_ord")))
plot(ggpredict(test, terms = c("age", "cannabism_group_ord")), show_ci = FALSE)


# Add predicted values to your data

dat_pred_nc_nacc1$predicted <-  predict(test, newdata = dat_pred_nc_nacc1, type = "response")




#If I want to color lines accordingly, I need to create average age by drinking-phase.
cutoff_age <- dat_pred_nc_nacc1 %>%
  filter(cannabis_phasem == "After") %>%
  group_by(subject) %>%
  summarize(min_after_age = min(age), .groups = "drop") %>%
  summarize(avg_cutoff = mean(min_after_age, na.rm = TRUE)) %>%
  pull(avg_cutoff)

dat_pred_nc_nacc1 <- dat_pred_nc_nacc1 %>%
  mutate(phase_age = case_when(
    cannabis_phasem == "Never" ~ "Never",
    age < cutoff_age ~ "Before",
    age >= cutoff_age ~ "After"
  ))

nacc1 <- ggplot(dat_pred_nc_nacc1, aes(x = age, y = predicted, group = subject, color = cannabis_phasem)) +
  #geom_point(alpha = 0.2, size = 0.05) +
  geom_line(alpha = 0.2, size = 0.1) +
      scale_color_manual(values = c(
    "After" = "#d95f02ff",
    "Before" = "black", 
    "Never" = "black"))   +   
    new_scale_color() +
  new_scale_fill() +
  geom_smooth(aes(group = phase_age, color = phase_age), size = 0.3, show.legend = FALSE) +
  labs(y = "", x = "") +
    scale_color_manual(values = c(
    "After" = "#d95f02ff",  
    "Before" = "black", 
    "Never" = "black"
  )) +
    facet_wrap(~cannabism_group, nrow = 1) + 
 # geom_vline(xintercept = cutoff_age, linetype = "dashed", size = 0.2) +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    legend.position = "none",
    strip.text = element_blank(),
    panel.background = element_blank(),   # remove gray panel
    panel.grid = element_blank(),         # remove grid lines
    plot.background = element_blank()     # remove outer gray
  )

nacc1

ggsave(filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/cannabis/monthly/NAcc_pred_ban_plot.pdf", 
      plot = nacc1,
       device = "pdf", 
       dpi = 300, 
      units = "mm",
       width = 90, 
       height = 60)

#Caudate


# Add predicted values to your data
dat_nc_long_ba_caud2 <- dat_nc_long_ba %>%
  filter(seed_cluster == "caudate_cluster2") 

#Seeing if additional grouping variable works
dat_nc_long_ba_caud2<- dat_nc_long_ba_caud2 %>%
  mutate(cannabism_group = ifelse(cannabis_phasem == "Never", "Never", "Initiated"),
         cannabism_group = factor(cannabism_group, levels = c("Initiated", "Never")))

#Oredered?
dat_nc_long_ba_caud2$cannabism_group_ord <- factor(dat_nc_long_ba_caud2$cannabism_group, 
                             levels = c("Never", "Initiated"), 
                             ordered = TRUE)

#Seeing if there are actual trajectory differences
test2 <- bam(mean_value ~ 1 + sex_fac  +  cannabism_group_ord +
              s(age, k = 5) +
              s(age, by = cannabism_group_ord, k = 5) +
              s(subject_fac, bs = 're') +
              s(age, subject_fac, bs = 're'), 
            data = dat_nc_long_ba_caud2, 
            na.action = na.exclude, 
            method = "fREML")

summary(test2)
draw(test2)
plot(ggpredict(test2, terms = c("age")))
plot(ggpredict(test2, terms = c("cannabism_group_ord")))
plot(ggpredict(test2, terms = c("age", "cannabism_group_ord")))

#Preds
dat_nc_long_ba_caud2$predicted <-  predict(test2, newdata = dat_nc_long_ba_caud2, type = "response")




#If I want to color lines accordingly, I need to create average age by drinking-phase.
cutoff_age <- dat_nc_long_ba_caud2 %>%
  filter(cannabis_phasem == "After") %>%
  group_by(subject) %>%
  summarize(min_after_age = min(age), .groups = "drop") %>%
  summarize(avg_cutoff = mean(min_after_age, na.rm = TRUE)) %>%
  pull(avg_cutoff)

dat_nc_long_ba_caud2 <- dat_nc_long_ba_caud2 %>%
  mutate(phase_age = case_when(
   cannabis_phasem == "Never" ~ "Never",
    age < cutoff_age ~ "Before",
    age >= cutoff_age ~ "After"
  ))

caud2 <- ggplot(dat_nc_long_ba_caud2, aes(x = age, y = predicted, group = subject, color = cannabis_phasem)) +
  #geom_point(alpha = 0.2, size = 0.05) +
  geom_line(alpha = 0.2, size = 0.1) +
      scale_color_manual(values = c(
    "After" = "#7500b3ff",
    "Before" = "black", 
    "Never" = "black"))   +   
    new_scale_color() +
  new_scale_fill() +
  geom_smooth(aes(group = phase_age, color = phase_age), size = 0.3, show.legend = FALSE) +
  labs(y = "", x = "") +
    scale_color_manual(values = c(
    "After" = "#7500b3ff",  
    "Before" = "black", 
    "Never" = "black"
  )) +
    facet_wrap(~cannabism_group, nrow = 1) + 
 # geom_vline(xintercept = cutoff_age, linetype = "dashed", size = 0.2) +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.title.x = element_text(size = 6, margin = margin(t = 3)),
    axis.title.y = element_text(size = 6, margin = margin(r = 3)),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    axis.line = element_line(size = .2),
    axis.ticks = element_line(size = .2),
    legend.position = "none",
    strip.text = element_blank(),
    panel.background = element_blank(),   # remove gray panel
    panel.grid = element_blank(),         # remove grid lines
    plot.background = element_blank()     # remove outer gray
  )

caud2

ggsave(filename = "H:/Dan/ncanda/R/figures/manuscript_figs/Supplementary_nicotine_cannabis/cannabis/monthly/caud_pred_ban_plot.pdf", 
      plot = caud2,
       device = "pdf", 
       dpi = 300, 
      units = "mm",
       width = 90, 
       height = 60)
```

