---
title: "ncanda_ASEBA_cleaning"
author: "Daniel Petrie"
date: "2024-08-26"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r Global}
#Packages
library(tidyverse) #Wrangling
library(dplyr) #Wrangling
library(stringr) #For str_replace
library(ggplot2)
library(psych)
library(BlandAltmanLeh)
#WD. Change whether through H: or on different sever
setwd("H:/Dan/ncanda/R")

#ASEBA data. I want OCD and stress variables
#Baseline through year 8
cbcl0_7 <- read.csv("H:/Raw/Datasets/NCANDA/followup_7y/NCANDA_RELEASE_7Y_REDCAP_MEASUREMENTS_V01/summaries/additional/cbc.csv", header = TRUE)
asr0_7 <- read.csv("H:/Raw/Datasets/NCANDA/followup_7y/NCANDA_RELEASE_7Y_REDCAP_MEASUREMENTS_V01/summaries/additional/asr.csv", header = TRUE)
ysr0_7 <- read.csv("H:/Raw/Datasets/NCANDA/followup_7y/NCANDA_RELEASE_7Y_REDCAP_MEASUREMENTS_V01/summaries/additional/ysr.csv", header = TRUE)
##Year 9
#I believe only ASR is collected. If youngest subject was 12, then they should be adults by now...
asr8 <- read.csv("H:/Raw/Datasets/NCANDA/followup_8y/NCANDA_RELEASE_8Y_REDCAP_MEASUREMENTS_V01/summaries/additional/asr.csv", header = TRUE)
```

Compulsions were assessed using three measures from the Achenback System of Empirically based Assessments (ASEBA; Achenbach & Rescorla (2001); Achenback & Rescorla (2003)):

Child Behavior Checklist (CBCL): Parent report for youth under age 18.
Youth Self-Report (YSR): Youth younger than 18.
Adult Self-Report (ASR): Youth older than 18.

Variable I want:
OCD (raw, t-score, percentile)


**Cleaning**

```{r Cleaning}
asr0_7_sub <- asr0_7 %>%
  select(subject, visit, asr_ocd_raw, asr_ocd_t, asr_ocd_pct)

asr8_sub <- asr8 %>%
  select(subject, visit, asr_ocd_raw, asr_ocd_t, asr_ocd_pct)

cbcl0_7_sub <- cbcl0_7 %>%
  select(subject, visit, cbcl_ocd_raw, cbcl_ocd_t, cbcl_ocd_pct)

ysr0_7_sub <- ysr0_7 %>%
  select(subject, visit, ysr_ocd_raw, ysr_ocd_t, ysr_ocd_pct)
```


I think my approach would be to show how similar YSR and CBCL are. If they are close, then I will just use YSR for youth and ASR for adults.

I'll create two files (eventually). First will be ASR/YSR for the main analytic data set. Second file will contain ASR/YSR/CBCL. This will be used to check how similar adult and youth reports are. This will be handled in a different script.


```{r}
#Full join to get a feel of the data
cbcl_ysr <- full_join(cbcl0_7_sub, ysr0_7_sub, by = c("subject", "visit"))

#Overall means
psych::describe(cbcl_ysr)
#Interesting, looks like t scores and % are a bit more similar.

#Correlation of raw scores
cor(cbcl_ysr$cbcl_ocd_raw, cbcl_ysr$ysr_ocd_raw, use = "complete.obs")
#cor(cbcl_ysr$cbcl_ocd_raw, cbcl_ysr$ysr_ocd_raw, use = "complete.obs", method = "spearman")

# Distributions, bivariate associations, and correlation
ggpairs(cbcl_ysr[,c("cbcl_ocd_raw","ysr_ocd_raw")])
```

Compute difference (youth - parent) score and see what that looks like. 

If +, youth scores are larger than parent scores, if -, than youth scores are smaller than parents.

If centered around 0, then that would imply that scores are reasonable.

```{r Difference score test}
#Difference score check.
cbcl_ysr$ocd_raw_diff <- cbcl_ysr$ysr_ocd_raw - cbcl_ysr$cbcl_ocd_raw
psych::describe(cbcl_ysr$ocd_raw_diff)

#Plotting difference scores as histogram
ggplot(cbcl_ysr, aes(x = ocd_raw_diff)) +
  geom_histogram(binwidth = 1, fill = "lightblue", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Difference Scores",
       x = "Difference Score (Child - Parent)",
       y = "Frequency") +
  theme_minimal()

#Plotting difference scores as boxplot
ggplot(cbcl_ysr, aes(y = ocd_raw_diff)) +
  geom_boxplot(fill = "orange", color = "black") +
  labs(title = "Boxplot of Difference Scores",
       y = "Difference Score (Child - Parent)") +
  theme_minimal()

#Plotting difference scores as scatterplot
ggplot(cbcl_ysr, aes(x = cbcl_ocd_raw, y = ysr_ocd_raw)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Child vs. Parent OCD Symptom Scores",
       x = "Parent Score",
       y = "Child Score") +
  theme_minimal()
```

T-test to look at difference

```{r t-test}
t.test(cbcl_ysr$cbcl_ocd_raw, cbcl_ysr$ysr_ocd_raw, paired = TRUE)
```

After doing all of this, I think the best approach is to use the YSR in the main analyses, then do a supplementary analysis with the cbcl scores as a replacement. I can also see what other OCD papers using the ASEBA have done for inspiration.

Next, I will combine YSR and ASR into one file and save.

```{r Merging ASR and YSR}
# YSR
ysr0_7_sub_merge <- ysr0_7_sub %>%
  mutate(aseba = "ysr") %>%
  #Rename columns by removing the "ysr_" prefix
  rename_with(~ gsub("^ysr_", "", .), starts_with("ysr_"))

# ASR 0-7
asr0_7_sub_merge <- asr0_7_sub %>%
  mutate(aseba = "asr") %>%
  #Rename columns by removing the "ysr_" prefix
  rename_with(~ gsub("^asr_", "", .), starts_with("asr_"))

# ASR 8
asr8_sub_merge <- asr8_sub %>%
  mutate(aseba = "asr") %>%
  #Rename columns by removing the "ysr_" prefix
  rename_with(~ gsub("^asr_", "", .), starts_with("asr_"))

#bind_rows and arrange by id and visit
ysr_asr <- bind_rows(ysr0_7_sub_merge, asr0_7_sub_merge, asr8_sub_merge) %>%
  arrange(subject, visit) %>%
   mutate(visit_numeric = case_when(
    visit == "baseline" ~ 0,
    visit == "followup_1y" ~ 1,
    visit == "followup_2y" ~ 2,
    visit == "followup_3y" ~ 3,
    visit == "followup_4y" ~ 4,
    visit == "followup_5y" ~ 5,
    visit == "followup_6y" ~ 6,
    visit == "followup_7y" ~ 7,
    visit == "followup_8y" ~ 8,
    TRUE ~ NA_real_  # Handle any unexpected visit values
  ))

#making visit numeric

```


Plotting some trajectories

```{r}
ggplot(ysr_asr %>% group_by(subject) %>% filter(n() > 7), 
       aes(x = visit_numeric, y = ocd_raw, group = subject)) +
  geom_line() +
  geom_point() +
  labs(title = "Longitudinal OCD Raw Scores",
       x = "Visit",
       y = "OCD Raw Score") +
  theme_minimal() +
  theme(legend.position = "none")
```


Looking at ICCs

```{r}
model_0 <- lmer(ocd_raw ~ 1 + (1|subject), data = ysr_asr, na.action = na.exclude)
summary(model_0)

icc(model_0)
#52% variability between people, and 48% variability within person. Great amount of within-person varaince to model!
ysr_asr$ocd_raw_pred <- predict(model_0)
ysr_asr$ocd_raw_resid<- residuals(model_0)

#plotting PREDICTED intraindividual change
ggplot(data = ysr_asr, aes(x = visit_numeric, y = ocd_raw_pred, group = subject)) +
  ggtitle("Unconditional Means Model") +
  geom_line() +
  xlab("Visit") + 
  ylab("PREDICTED OCD Raw Score") +
  geom_smooth(method = "lm", group = 1)

ggplot(data = ysr_asr, aes(x = visit_numeric, y = ocd_raw_resid, group = subject)) +
  ggtitle("Unconditional Means Model (residuals)") +
  geom_line() +
  xlab("Visit") + 
  ylab("RESIDUALS OCD Raw Score") +
  geom_smooth(method = "lm", group = 1)

```


Seeing if growth is a thing.

```{r}
model_1 <- lmer(ocd_raw ~ 1 + visit_numeric + (1|subject), data = ysr_asr, na.action = na.exclude)
summary(model_1)

#52% variability between people, and 48% variability within person. Great amount of within-person varaince to model!
ysr_asr$ocd_raw_pred1 <- predict(model_1)
ysr_asr$ocd_raw_resid1 <- residuals(model_1)

#plotting PREDICTED intraindividual change
ggplot(data = ysr_asr, aes(x = visit_numeric, y = ocd_raw_pred1, group = subject)) +
  ggtitle("Fixed Linear, Random Intercept") +
  geom_line() +
  xlab("Visit") + 
  ylab("PREDICTED OCD Raw Score") +
  geom_smooth(method = "lm", group = 1)

ggplot(data = ysr_asr, aes(x = visit_numeric, y = ocd_raw_resid1, group = subject)) +
  ggtitle("Fixed Linear, Random Intercept (residuals)") +
  geom_line() +
  xlab("Visit") + 
  ylab("RESIDUALS OCD Raw Score") +
  geom_smooth(method = "lm", group = 1)

```


Seeing what random intercepts look like

Seeing if growth is a thing.

```{r}
model_2 <- lmer(ocd_raw ~ 1 + visit_numeric + (1 + visit_numeric|subject), data = ysr_asr, na.action = na.exclude)
summary(model_2)

ysr_asr$ocd_raw_pred2 <- predict(model_2)
ysr_asr$ocd_raw_resid2 <- residuals(model_2)

#plotting PREDICTED intraindividual change
ggplot(data = ysr_asr, aes(x = visit_numeric, y = ocd_raw_pred2, group = subject)) +
  ggtitle("Fixed Linear, Random Linear") +
  geom_line() +
  xlab("Visit") + 
  ylab("PREDICTED OCD Raw Score") +
  geom_smooth(method = "lm", group = 1)

ggplot(data = ysr_asr, aes(x = visit_numeric, y = ocd_raw_resid2, group = subject)) +
  ggtitle("Fixed Linear, Random Intercept (residuals)") +
  geom_line() +
  xlab("Visit") + 
  ylab("RESIDUALS OCD Raw Score") +
  geom_smooth(method = "lm", group = 1)
```

We can also compare models to see if adding random slope (juice) is needed (worth the squeeze)

```{r}
anova(model_1, model_2)
```

Results from the anova test suggest that the more complex model with random slopes is a better fit.

Save csv in ./clean_data

```{r}
#write.csv(ysr_asr, file = "./clean_data/ysr_asr0_8.csv", row.names = FALSE)
```