---
title: "ncanda_cleaning"
author: "Daniel Petrie"
date: "2024-08-20"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r Global}
#Packages
library(tidyverse) #Wrangling
library(dplyr) #Wrangling
library(stringr) #For str_replace
library(ggplot2)
library(psych)
library(see) #theme_modern()
#WD. Change whether through H: or on different sever
setwd("H:/Dan/ncanda/R")

#CDDR data. I think I want to use the "youth report" version for now because it has ages.
#Baseline through year 8
cddr0_7 <- read.csv("H:/Raw/Datasets/NCANDA/followup_7y/NCANDA_RELEASE_7Y_REDCAP_MEASUREMENTS_V01/summaries/redcap/youthreport1_cddr.csv", header = TRUE)
#Year 9
cddr8 <- read.csv("H:/Raw/Datasets/NCANDA/followup_8y/NCANDA_RELEASE_8Y_REDCAP_MEASUREMENTS_V01/summaries/redcap/youthreport1_cddr.csv", header = TRUE)
```

Cleaning 
add youthreport1_cddr15

```{r CDDR Cleaining}
#Vector of alcohol variable names
subset_var <- c("subject", "visit", "youthreport1_age", "youthreport1_cddr16",
                "youthreport1_cddr17", "youthreport1_cddr18", "youthreport1_cddr19", "youthreport1_cddr21",
                "youthreport1_cddr22", "youthreport1_cddr23", "youthreport1_cddr26", "youthreport1_cddr27",
                "youthreport1_cddr30", "youthreport1_cddr31")

#Binding data, arrange by visit, select alcohol variables, rename.
cddr0_8 <- bind_rows(cddr0_7, cddr8) %>% 
  arrange(subject, visit) %>%
  select(all_of(subset_var)) %>%
  rename_with(~ str_replace(., "^youthreport1_", ""), starts_with("youthreport1_"))

#Descriptives
psych::describe(cddr0_8)
```


**CDDR variable descriptions**
11 variables total.

cddr16: How old were you when you first drank ALCOHOL regularly? (at least once a week; beer, wine, or hard liquor)  

cddr17: When was the last time you drank ALCOHOL? (0=today, 1=yesterday, etc.)

cddr18: How many ALCOHOLIC DRINKS did you drink on that occasion?

cddr19: In your lifetime, how many days did you drink ALCOHOL (beer, wine, hard liquor)?

cddr21: During the past year, (so out of 365 days), how many days did you drink ALCOHOL (beer, wine, hard liquor)? (0-365)

cddr22: During the last 30 days, how many days did you drink ALCOHOL? (0-30)

cddr23: Over the last month, in the average 24-hour period you were drinking, how many ALCOHOLIC DRINKS did you have?

cddr26: During the past year (or 365 days), in the average 24-hour period you were drinking how many total drinks did you have? 

cddr27: Have you ever consumed 4+ (females) / 5+ (males) drinks within an occasion? (y/n)

cddr30: During the past year, how many times have you consumed 4+ (females) / 5+ (males) drinks within an occasion? (0-365) 

cddr31: During the last 30 days, how many times have you consumed 4+ (females) / 5+ (males) drinks within an occasion? (0-30)


Recoding some NAs to 0's to reflect an absence of drinking behavior and not a true missing value. DOing this for cddr21/22/30/31

```{r}
# Recode past year drinking days to 0 if participants skipped out of CDDR items due to no history of use
cddr0_8$cddr21 <- ifelse(is.na(cddr0_8$cddr21),
                                 ifelse(is.na(cddr0_8$age),NA,0),cddr0_8$cddr21)

# Recode past month drinking days to 0 if participants skipped out of CDDR items due to no history of use
cddr0_8$cddr22 <- ifelse(is.na(cddr0_8$cddr22),
                                 ifelse(is.na(cddr0_8$age),NA,0),cddr0_8$cddr22)

# Recode PastYear Binges to 0 if participants skipped out of CDDR items due to no history of use
cddr0_8$cddr30 <- ifelse(is.na(cddr0_8$cddr30),
                                 ifelse(is.na(cddr0_8$age),NA,0),cddr0_8$cddr30)

# Recode Binge Past Month
cddr0_8$cddr31 <- ifelse(is.na(cddr0_8$cddr31),
                                 ifelse(is.na(cddr0_8$age),NA,0),cddr0_8$cddr31)
```



Composite alcohol scoring. Kiss 

```{r}

```


Some plots

```{r}
response_columns <- c("cddr16", "cddr17", "cddr18", "cddr19", 
"cddr21", "cddr22", "cddr23", "cddr26", "cddr27", "cddr30", "cddr31")


cddr_plots <- list()


# Loop through each response column and fit the model
for (response in response_columns) {
  
  # Plot
  cddr_plots[[response]] <- ggplot(cddr0_8, aes(x = age, 
                                                y = .data[[response]], 
                                                group = subject)) + 
    geom_line(alpha = 0.2) +
    geom_smooth(aes(group = 1), 
                method = "loess", 
                color = "blue") +
    labs(title = paste("Plot for", response), 
         x = "Age", 
         y = response) +
  theme_modern()
}
```

Looking at mean alcohol use past year across time points.

```{r}
cddr0_8 %>%
  group_by(visit) %>%
  summarize(mean_cddr21 = mean(cddr21, na.rm = TRUE))
```


Write csv in clean_data

```{r}
write.csv(cddr0_8, file = "./clean_data/cddr0_8.csv", row.names = FALSE)
```

