---
title: "ncanda_motion"
author: "Daniel Petrie"
date: "2024-08-26"
output: html_document
editor_options: 
  chunk_output_type: console
---

Global

```{r}
#Packages
library(tidyverse) #Wrangling
library(dplyr) #Wrangling
library(stringr) #For str_replace
library(ggplot2)
library(psych)
library(lme4)
library(lmerTest)
library(GGally)
#library(pscl)
library(glmmTMB)
#WD. Change whether through H: or on different sever
setwd("H:/Dan/ncanda/")

#Data (for ages)
cddr0_8 <- read.csv("./R/clean_data/cddr0_8.csv", header = TRUE)

#Motion
ge_motion <- read.table("./motion_txt/ncanda_ge_motion.txt", header = FALSE, sep = "")
siemens_motion <- read.delim("./motion_txt/ncanda_siemens_motion.txt", header = FALSE, sep = "")
```

Combine and clean

```{r}
#Adding scanner column just in case.
ge_motion$scanner <- "ge"
siemens_motion$scanner <- "siemens"

#Creating the motion file and renaming.
motion <- 
  bind_rows(ge_motion, siemens_motion) %>%
  select(-V1) %>%
  rename(id = V2,
         count = V3)


cddr0_8_merge <- cddr0_8 %>%
  mutate(visitnum = case_when(
    visit == "baseline" ~ "0",
    visit == "followup_1y" ~ "1",
    visit == "followup_2y" ~ "2",
    visit == "followup_3y" ~ "3",
    visit == "followup_4y" ~ "4",
    visit == "followup_5y" ~ "5",
    visit == "followup_6y" ~ "6",
    visit == "followup_7y" ~ "7",
    visit == "followup_8y" ~ "8")) %>%
  mutate(id = paste0(gsub("NCANDA_", "", subject), "_", visitnum)) %>%
  select(id, age, visit)

#Merging
motion_age <- left_join(motion, cddr0_8_merge, by = "id")
```

Ok. Clean enough. Plotting motion over age and binned motion

```{r}
#Amount of volumes that are acceptable
270*.3
#81

#Gettting a sense of motion for whole sample
ggplot(motion_age, aes(x = count)) +
  geom_histogram(binwidth = 1, 
                 color = "black", 
                 fill = "lightblue") +
  labs(x = "Count", 
       y = "Frequency", 
       title = "Frequency of Count Values") +
  theme_modern() +
  geom_vline(xintercept = 81, 
             linetype = "dashed", 
             color = "orange", 
             size = 1)

ggplot(motion_age, aes(x = age, y = count)) +
  geom_point() +
  geom_hline(yintercept = 81, linetype = "dashed", color = "orange", size = 1) +
  labs(x = "Age", 
       y = "Censored TRs", 
       title = "Motion across age") +
  theme_modern() +
  geom_smooth(method = "lm")


ggplot(motion_age, aes(x = visit, y = count)) +
  geom_point() +
  geom_hline(yintercept = 81, linetype = "dashed", color = "orange", size = 1) +
  labs(x = "Age", 
       y = "Censored TRs", 
       title = "Motion across visit") +
  theme_modern()

#Age effect on motion
summary(lm(count~1 + age, data = motion_age))

#A 1 unit increase in age is associated with -0.3 unit decrease in motion. In other words, younger subjects have more motion.
```

Now I need to flag any instances where the count of censored TRs > 81 and get a rough breakdown of their ages and what scanner they used.

```{r}
motion_age_removed <- motion_age %>%
  mutate(remove = ifelse(count >= 81, TRUE, FALSE)) %>%
  filter(remove == TRUE)

145/4505 # 3% of scans (145) have high motion.

ggplot(motion_age_removed, aes(x = age, y = count)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Age", 
       y = "Censored TRs", 
       title = "Motion across age") +
  theme_modern()

ggplot(motion_age_removed, aes(x = visit, y = count)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Visit", 
       y = "Censored TRs", 
       title = "Motion across age") +
  theme_modern()

#No effect of age on motion
summary(lm(count~1 + age, data = motion_age_removed))
```

Appears to even be a slight increase in censored TRs with age, but basically flat.

For main cleaning script, I should add a line of code removing subjects with bad imaging from this script.

Creating .csv to filter subject with high motion.

```{r}
write.csv(motion_age_removed, file = "./R/clean_data/high_motion_subjects.csv", row.names = FALSE)
```

