---
title: "ncanda_seed_cleaning"
author: "Daniel Petrie"
date: "2024-11-11"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library("ggplot2") #For plotting
library("GGally") #ggpairs()
library("tidyverse") #Wranglin
library("dplyr") #Wranglin
library("ggseg")
library("ggsegGordon")
library("lme4")
library(lmerTest)
setwd("H:/Dan/ncanda/R")


#Putamen
put_seed <- read.delim("H:/Dan/ncanda/rest/putamen_seed_gordon_harox.tsv", header = TRUE)

#Caudate
caud_seed <- read.delim("H:/Dan/ncanda/rest/caudate_seed_gordon_harox.tsv", header = TRUE)

#Accumbens
nacc_seed <- read.delim("H:/Dan/ncanda/rest/accumbens_seed_gordon_harox.tsv", header = TRUE)

#Gordon parcels
parcels <- read.csv("Parcels.csv", header = TRUE)
```

Four data frames based on seeded connectivity analyses:
1.) Putamen seed
2.) Caudate seed
3.) Nacc seed
4.) Network-level from all seeds

It might make the most sense to keep separate for now and treat like three different data sets. I will start with just the putamen one for analyses for NCANDA abstract. For the manuscript, I'll either combine or use as specificity type of thing.


```{r Cleaning}
#Putamen
put_seed_clean <- put_seed %>%
  mutate(
    subject = str_extract(name, "S\\d+"),
    visit_numeric = str_extract(name, "(?<=_)(\\d+)(?=_seed)") %>% 
      as.numeric()
  ) %>%
  select(-contains("count"), -X, -X.1, -name) %>%
  select(subject, visit_numeric, everything())

#Caudate
caud_seed_clean <- caud_seed %>%
  mutate(
    subject = str_extract(name, "S\\d+"),
    visit_numeric = str_extract(name, "(?<=_)(\\d+)(?=_seed)") %>% 
      as.numeric()
  ) %>%
  select(-contains("count"), -X, -X.1, -name) %>%
  select(subject, visit_numeric, everything())

#Accumbens
nacc_seed_clean <- nacc_seed %>%
  mutate(
    subject = str_extract(name, "S\\d+"),
    visit_numeric = str_extract(name, "(?<=_)(\\d+)(?=_seed)") %>% 
      as.numeric()
  ) %>%
  select(-contains("count"), -X, -X.1, -name) %>%
  select(subject, visit_numeric, everything())

#Parcel labels
#parcels_clean <- parcels %>%
#  mutate(label = paste(Hem, Community, ParcelID, sep = "_"))

parcels_clean <- parcels %>%
   mutate(
    Community = case_when(
      Community == "CinguloParietal" ~ "MedialParietal",
      Community == "RetrosplenialTemporal" ~ "ParietoOccip",
      TRUE ~ Community  # Keep other values unchanged
    )) %>%
  group_by(Community) %>%
  mutate(region = paste0(Community, "_", row_number())) %>%
  ungroup()

#CinguloParietal changed to MedialParietal
#RetrosplenialTemporal changed to ParietoOccip
```

Note. For plotting with ggseg() I need to figure out some additional labeling. I have them numbered sequentially, but I really probably need them numbered according to what function network they belong to.


This was afni command that created HarOx subcort.

3dcalc -a HarOx-sub-1mm+tlrc. -expr '1*equals(a,4) + 2*equals(a,5) + 3*equals(a,6) + 4*equals(a,7) + 5*equals(a,10) + 6*equals(a,11) + 7*equals(a,15) + 8*equals(a,16) + 9*equals(a,17) + 10*equals(a,18) + 11*equals(a,20) + 12*equals(a,21)' -prefix HarOx-sub-1mm-mod+tlrc

This was afni command to make Gordon_HarOx

3dcalc -a Parcels_MNI_111_warped.nii.gz -b HarOx-sub-1mm-mod+tlrc. -expr 'a + (333*ispositive(b-0.5) + b)*equals(a,0)' -prefix Gordon_HarOx_1mm+tlrc

These are labels from HarvardOxford-Subcortical.xml in FSL directory.

ROI           HarOx Label # (AFNI new atlas #)
thalamus = 4 (1),15 (7)
amygdala = 10 (5),20 (11)
nacc = 11 (6),21 (12)
putamen = 6 (3),17 (9)
caudate = 5 (2),16 (8)
pallidum = 7 (4),18 (10)

```{r Replacing column names Putamen}
# Create mapping of ParcelID to label
label_map <- parcels_clean %>% 
  select(ParcelID, region) %>% 
  deframe()

# Extract the numeric part from each target_df column and replace if it exists in the label_map
new_colnames <- colnames(put_seed_clean) %>% 
  # Apply the replacement based on ParcelID
  purrr::map_chr(~ ifelse(str_detect(.x, "NZMean_\\d+"),
                          label_map[str_extract(.x, "\\d+")], # Replace with matching label if available
                          .x))  # Otherwise, keep the original name

# Set the new column names
colnames(put_seed_clean) <- new_colnames

# View the updated target_df
head(put_seed_clean)

# Creating vector of names for subcortical based on indexing above.
subcort <- c("L_Thalamus", "L_Caudate", "L_Putamen",
             "L_Pallidum", "L_Amygdala", "L_Nacc",
             "R_Thalamus", "R_Caudate", "R_Putamen",
             "R_Pallidum", "R_Amygdala", "R_Nacc")

# Verify that the subcort vector has the same length as the number of NA columns
if(length(subcort) == length(336:347)) {
  # Assign subcort vector names to the NA columns in positions 336 to 347
  names(put_seed_clean)[336:347] <- subcort
} else {
  stop("The length of 'subcort' does not match the number of columns to rename.")
}

colnames(put_seed_clean[336:347])
```

```{r Replacing column names Caudate}
# Create mapping of ParcelID to label
label_map <- parcels_clean %>% 
  select(ParcelID, region) %>% 
  deframe()

# Extract the numeric part from each target_df column and replace if it exists in the label_map
new_colnames <- colnames(caud_seed_clean) %>% 
  # Apply the replacement based on ParcelID
  purrr::map_chr(~ ifelse(str_detect(.x, "NZMean_\\d+"),
                          label_map[str_extract(.x, "\\d+")], # Replace with matching label if available
                          .x))  # Otherwise, keep the original name

# Set the new column names
colnames(caud_seed_clean) <- new_colnames

# View the updated target_df
head(caud_seed_clean)

# Creating vector of names for subcortical based on indexing above.
subcort <- c("L_Thalamus", "L_Caudate", "L_Putamen",
             "L_Pallidum", "L_Amygdala", "L_Nacc",
             "R_Thalamus", "R_Caudate", "R_Putamen",
             "R_Pallidum", "R_Amygdala", "R_Nacc")

# Verify that the subcort vector has the same length as the number of NA columns
if(length(subcort) == length(336:347)) {
  # Assign subcort vector names to the NA columns in positions 336 to 347
  names(caud_seed_clean)[336:347] <- subcort
} else {
  stop("The length of 'subcort' does not match the number of columns to rename.")
}

colnames(caud_seed_clean[336:347])
```

```{r Replacing column names accumbens}
# Create mapping of ParcelID to label
label_map <- parcels_clean %>% 
  select(ParcelID, region) %>% 
  deframe()

# Extract the numeric part from each target_df column and replace if it exists in the label_map
new_colnames <- colnames(nacc_seed_clean) %>% 
  # Apply the replacement based on ParcelID
  purrr::map_chr(~ ifelse(str_detect(.x, "NZMean_\\d+"),
                          label_map[str_extract(.x, "\\d+")], # Replace with matching label if available
                          .x))  # Otherwise, keep the original name

# Set the new column names
colnames(nacc_seed_clean) <- new_colnames

# View the updated target_df
head(nacc_seed_clean)

# Creating vector of names for subcortical based on indexing above.
subcort <- c("L_Thalamus", "L_Caudate", "L_Putamen",
             "L_Pallidum", "L_Amygdala", "L_Nacc",
             "R_Thalamus", "R_Caudate", "R_Putamen",
             "R_Pallidum", "R_Amygdala", "R_Nacc")

# Verify that the subcort vector has the same length as the number of NA columns
if(length(subcort) == length(336:347)) {
  # Assign subcort vector names to the NA columns in positions 336 to 347
  names(nacc_seed_clean)[336:347] <- subcort
} else {
  stop("The length of 'subcort' does not match the number of columns to rename.")
}

colnames(nacc_seed_clean[336:347])
```

Creating network level data. This should be expanded to include the other seed ROIs.

```{r}
#putamen
put_seed_clean_net <- put_seed_clean %>%
  rowwise() %>%
  mutate(
    Auditory_putamen = mean(c_across(contains("Auditory")), na.rm = TRUE),
    CinguloOperc_putamen = mean(c_across(contains("CinguloOperc")), na.rm = TRUE),
    MedialParietal_putamen = mean(c_across(contains("MedialParietal")), na.rm = TRUE),
    Default_putamen = mean(c_across(contains("Default")), na.rm = TRUE),
    DorsalAttn_putamen = mean(c_across(contains("DorsalAttn")), na.rm = TRUE),
    FrontoParietal_putamen = mean(c_across(contains("FrontoParietal")), na.rm = TRUE),
    None_putamen = mean(c_across(contains("None")), na.rm = TRUE),
    ParietoOccip_putamen = mean(c_across(contains("ParietoOccip")), na.rm = TRUE),
    Salience_putamen = mean(c_across(contains("Salience")), na.rm = TRUE),
    SMhand_putamen = mean(c_across(contains("SMhand")), na.rm = TRUE),
    SMmouth_putamen = mean(c_across(contains("SMmouth")), na.rm = TRUE),
    VentralAttn_putamen = mean(c_across(contains("VentralAttn")), na.rm = TRUE),
    Visual_putamen = mean(c_across(contains("Visual")), na.rm = TRUE)
  ) %>%
  ungroup()

#caudate
caud_seed_clean_net <- caud_seed_clean %>%
  rowwise() %>%
  mutate(
    Auditory_caudate = mean(c_across(contains("Auditory")), na.rm = TRUE),
    CinguloOperc_caudate = mean(c_across(contains("CinguloOperc")), na.rm = TRUE),
    MedialParietal_caudate = mean(c_across(contains("MedialParietal")), na.rm = TRUE),
    Default_caudate = mean(c_across(contains("Default")), na.rm = TRUE),
    DorsalAttn_caudate = mean(c_across(contains("DorsalAttn")), na.rm = TRUE),
    FrontoParietal_caudate = mean(c_across(contains("FrontoParietal")), na.rm = TRUE),
    None_caudate = mean(c_across(contains("None")), na.rm = TRUE),
    ParietoOccip_caudate = mean(c_across(contains("ParietoOccip")), na.rm = TRUE),
    Salience_caudate = mean(c_across(contains("Salience")), na.rm = TRUE),
    SMhand_caudate = mean(c_across(contains("SMhand")), na.rm = TRUE),
    SMmouth_caudate = mean(c_across(contains("SMmouth")), na.rm = TRUE),
    VentralAttn_caudate = mean(c_across(contains("VentralAttn")), na.rm = TRUE),
    Visual_caudate = mean(c_across(contains("Visual")), na.rm = TRUE)
  ) %>%
  ungroup()

#accumbens
nacc_seed_clean_net <- nacc_seed_clean %>%
  rowwise() %>%
  mutate(
    Auditory_accumbens = mean(c_across(contains("Auditory")), na.rm = TRUE),
    CinguloOperc_accumbens = mean(c_across(contains("CinguloOperc")), na.rm = TRUE),
    MedialParietal_accumbens = mean(c_across(contains("MedialParietal")), na.rm = TRUE),
    Default_accumbens = mean(c_across(contains("Default")), na.rm = TRUE),
    DorsalAttn_accumbens = mean(c_across(contains("DorsalAttn")), na.rm = TRUE),
    FrontoParietal_accumbens = mean(c_across(contains("FrontoParietal")), na.rm = TRUE),
    None_accumbens = mean(c_across(contains("None")), na.rm = TRUE),
    ParietoOccip_accumbens = mean(c_across(contains("ParietoOccip")), na.rm = TRUE),
    Salience_accumbens = mean(c_across(contains("Salience")), na.rm = TRUE),
    SMhand_accumbens = mean(c_across(contains("SMhand")), na.rm = TRUE),
    SMmouth_accumbens = mean(c_across(contains("SMmouth")), na.rm = TRUE),
    VentralAttn_accumbens = mean(c_across(contains("VentralAttn")), na.rm = TRUE),
    Visual_accumbens = mean(c_across(contains("Visual")), na.rm = TRUE)
  ) %>%
  ungroup()
```

Combining the network-level data into a separate data frame.

```{r}
put_seed_clean_sub <- put_seed_clean_net %>%
  select(subject, visit_numeric, Auditory_putamen, CinguloOperc_putamen, MedialParietal_putamen, Default_putamen, DorsalAttn_putamen, FrontoParietal_putamen, None_putamen, ParietoOccip_putamen, Salience_putamen, SMhand_putamen, SMmouth_putamen, VentralAttn_putamen, Visual_putamen)

caud_seed_clean_sub <- caud_seed_clean_net %>%
  select(subject, visit_numeric, Auditory_caudate, CinguloOperc_caudate, MedialParietal_caudate, Default_caudate, DorsalAttn_caudate, FrontoParietal_caudate, None_caudate, ParietoOccip_caudate, Salience_caudate, SMhand_caudate, SMmouth_caudate, VentralAttn_caudate, Visual_caudate)

nacc_seed_clean_sub <- nacc_seed_clean_net %>%
  select(subject, visit_numeric, Auditory_accumbens, CinguloOperc_accumbens, MedialParietal_accumbens, Default_accumbens, DorsalAttn_accumbens, FrontoParietal_accumbens, None_accumbens, ParietoOccip_accumbens, Salience_accumbens, SMhand_accumbens, SMmouth_accumbens, VentralAttn_accumbens, Visual_accumbens)

#Combining
combined_seed_networks <- merge(put_seed_clean_sub, caud_seed_clean_sub, by = c("subject", "visit_numeric"))
combined_seed_networks <- merge(combined_seed_networks, nacc_seed_clean_sub, by = c("subject", "visit_numeric"))
```


Saving. This should be done across caudate and accumbens at ROI level. Could also be worth it to create a data frame where it is just the gordon networks for each seed region. For example, there are 13 gordon networks, across three ROIs. So auditory_putamen, auditory_caudate, auditory_accumbens for the main data set. For now I will continue to just focus on the putamen seeded network.

```{r Saving}
write.csv(put_seed_clean, file = "clean_data/putamen_seed_clean.csv", row.names = FALSE)
write.csv(caud_seed_clean, file = "clean_data/caudate_seed_clean.csv", row.names = FALSE)
write.csv(nacc_seed_clean, file = "clean_data/nacc_seed_clean.csv", row.names = FALSE)
write.csv(combined_seed_networks, file = "clean_data/combined_seed_networks_clean.csv", row.names = FALSE)
```